### 2017-03-17 paper/lipics-v2016-sample-article.tex
Work - Paper style and stuff

\documentclass[a4paper,UKenglish]{lipics-v2016}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
% for section-numbered lemmas etc., use "numberwithinsect"

\usepackage{microtype}%if unwanted, comment out or use option "draft"

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the recommended bibstyle

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{A Sample Article for the LIPIcs series\footnote{This work was partially supported by someone.}}
\titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{John Q. Open}
\author[2]{Joan R. Access}
\affil[1]{Dummy University Computing Laboratory, Address/City, Country\\
  \texttt{open@dummyuniversity.org}}
\affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
  \texttt{access@dummycollege.org}}
\authorrunning{J.\,Q. Open and J.\,R. Access} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{John Q. Open and Joan R. Access}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. 
 \end{abstract}

\section{Typesetting instructions -- please read carefully}
Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
\begin{itemize}
\item Use pdflatex and an up-to-date LaTeX system.
\item Use further LaTeX packages only if required. Avoid usage of packages like \verb+enumitem+, \verb+enumerate+, \verb+cleverref+. Keep it simple, i.e. use as few additional packages as possible.
\item Add custom made macros carefully and only those which are needed in the article (i.e., do not simply add your convolute of macros collected over the years).
\item Do not use a different main font. For example, the usage of the \verb+times+-package is forbidden.
\item Provide full author names (especially with regard to the first name) in the \verb+\author+ macro and in the \verb+\Copyright+ macro.
\item Fill out the \verb+\subjclass+ and \verb+\keywords+ macros. For the \verb+\subjclass+, please refer to the ACM classification at \url{http://www.acm.org/about/class/ccs98-html}.
\item Take care of suitable linebreaks and pagebreaks. No overfull \verb+\hboxes+ should occur in the warnings log.
\item Provide suitable graphics of at least 300dpi (preferrably in pdf format).
\item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsection*+, \verb+\paragraph+, \verb+\subparagraph*+, ... ``Self-made'' sectioning commands (for example, \verb+\noindent{\bf My+ \verb+subparagraph.}+ will be removed and replaced by standard LIPIcs style sectioning commands.
\item Do not alter the spacing of the  \verb+lipics-v2016.cls+ style file. Such modifications will be removed.
\item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
\item Remove all comments, especially avoid commenting large text blocks and using \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions.
\item Keep the standard style (\verb+plainurl+) for the bibliography as provided by the\linebreak \verb+lipics-v2016.cls+ style file.
\item Use BibTex and provide exactly one BibTex file for your article. The BibTex file should contain only entries that are referenced in the article. Please make sure that there are no errors and warnings with the referenced BibTex entries.
\item Use a spellchecker to get rid of typos.
\item A manual for the LIPIcs style is available at \url{http://drops.dagstuhl.de/styles/lipics-v2016/lipics-v2016-authors/lipics-v2016-manual.pdf}.
\end{itemize}

\section{Lorem ipsum dolor sit amet}

Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. Donec eget odio et magna ullamcorper vehicula ut vitae libero. Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis. Pellentesque id ante erat. In hac habitasse platea dictumst. Curabitur a scelerisque odio. Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam. Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.

\begin{lemma}[Lorem ipsum]
\label{lemma:lorem}
Vestibulum sodales dolor et dui cursus iaculis. Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper. Proin facilisis gravida rutrum. Etiam sed sollicitudin lorem. Proin pellentesque risus at elit hendrerit pharetra. Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
\end{lemma}

\begin{proof}
Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.
\end{proof}

\begin{theorem}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
\label{theorem:curabitur}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\end{theorem}

\subsection{Curabitur dictum felis id sapien}

Curabitur dictum felis id sapien mollis ut venenatis tortor feugiat. Curabitur sed velit diam. Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu. Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet. Duis rutrum dapibus ornare. Nulla vehicula vulputate iaculis. Proin a consequat neque. Donec ut rutrum urna. Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nec faucibus leo. Cras ut nisl odio, non tincidunt lorem. Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat. Fusce risus libero, convallis at fermentum in, dignissim sed sem. Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}. Donec non suscipit lorem. Nam sit amet enim vitae nisl accumsan pretium.

\begin{lstlisting}[caption={Useless code},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
for i:=maxint to 0 do 
begin 
    j:=square(root(i));
end;
\end{lstlisting}

\subsection{Proin ac fermentum augue}

Proin ac fermentum augue. Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis. Nulla facilisi. Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac. Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna. Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui. Praesent a pretium quam. Mauris non mauris sit amet eros rutrum aliquam id ut sapien. Nulla aliquet fringilla sagittis. Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor. Nulla cursus aliquet enim. Cras sapien risus, accumsan eu cursus ut, commodo vel velit. Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel. Integer faucibus faucibus felis.

\begin{itemize}
\item Ut vitae diam augue. 
\item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
\item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
\end{itemize}

\section{Pellentesque quis tortor}

Nec urna malesuada sollicitudin. Nulla facilisi. Vivamus aliquam tempus ligula eget ornare. Praesent eget magna ut turpis mattis cursus. Aliquam vel condimentum orci. Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi. Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.

\begin{lemma} [Quisque blandit tempus nunc]
Sed interdum nisl pretium non. Mauris sodales consequat risus vel consectetur. Aliquam erat volutpat. Nunc sed sapien ligula. Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus. Nunc vestibulum nunc ac massa pretium pharetra. Nulla facilisis turpis id augue venenatis blandit. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
\end{lemma}

Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.

\subparagraph*{Acknowledgements.}

I want to thank \dots

\appendix
\section{Morbi eros magna}

Morbi eros magna, vestibulum non posuere non, porta eu quam. Maecenas vitae orci risus, eget imperdiet mauris. Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis. Sed condimentum convallis dolor, a dignissim est ultrices eu. Donec consectetur volutpat eros, et ornare dui ultricies id. Vivamus eu augue eget dolor euismod ultrices et sit amet nisi. Vivamus malesuada leo ac leo ullamcorper tempor. Donec justo mi, tempor vitae aliquet non, faucibus eu lacus. Donec dictum gravida neque, non porta turpis imperdiet eget. Curabitur quis euismod ligula.

%%
%% Bibliography
%%

%% Either use bibtex (recommended),

\bibliography{lipics-v2016-sample-article}

%% .. or use the thebibliography environment explicitely

\end{document}

### 2017-03-17 include/cuckoo_base.h

@@ -69,13 +69,26 @@ public:
     using value_type      = std::pair<const key_type, mapped_type>;
     using iterator        = IteratorBase<iterator_incr<Specialized_t> >;
     using const_iterator  = IteratorBase<iterator_incr<Specialized_t>, true>;
+    using size_type       = size_t;
+    using difference_type = std::ptrdiff_t;
+    // using hasher          = Hash;
+    // using key_equal       = KeyEqual
+    // using allocator_type  = Allocator
+    using reference       = value_type&;
+    using const_reference = const value_type&;
+    // using pointer         = std::allocator_traits<Allocator>::pointer;
+    // using const_pointer   = std::allocator_traits<Allocator>::const_pointer;
+    using insert_return_type   = std::pair<iterator, bool>;
 
+    using local_iterator       = void;
+    using const_local_iterator = void;
+    using node_type            = void;
 private:
     using  value_intern    = std::pair<key_type, mapped_type>;
 
 public:
     CuckooMultiBase(double size_constraint = 1.1,
-                    size_t dis_steps = 0, size_t seed = 0);
+                    size_type dis_steps = 0, size_type seed = 0);
     ~CuckooMultiBase() = default;
     CuckooMultiBase(const CuckooMultiBase&     ) = delete;
     CuckooMultiBase(      CuckooMultiBase&& rhs);
@@ -88,58 +101,70 @@ public:
     }
 
 private:
-    // members that should become private at some point ************************
-    size_t       n;
-    size_t       capacity;
-    size_t       grow_thresh;
+    // Members *****************************************************************
+    size_type    n;
+    size_type    capacity;
+    size_type    grow_thresh;
     double       alpha;
     Hasher_t     hasher;
     DisStrat_t   displacer;
     HistCount_t  hcounter;
-    static constexpr size_t bs = CuckooTraits<Specialized_t>::bs;
-    static constexpr size_t tl = CuckooTraits<Specialized_t>::tl;
-    static constexpr size_t nh = CuckooTraits<Specialized_t>::nh;
+    static constexpr size_type bs = CuckooTraits<Specialized_t>::bs;
+    static constexpr size_type tl = CuckooTraits<Specialized_t>::tl;
+    static constexpr size_type nh = CuckooTraits<Specialized_t>::nh;
 
 public:
[…]

### 2017-03-17 include/cuckoo_eg2l.h

@@ -123,7 +123,7 @@ public:
                         : std::make_pair(0,nullptr);
     }
 
-    iterator begin() const
+    iterator begin()
     {
         auto temp = make_iterator(&llt[0][0].elements[0]);
         if (! temp->first) temp++;

### 2017-03-17 include/cuckoo_homogeneous2l.h

@@ -22,8 +22,8 @@ private:
     friend iterator_incr<This_t>;
 
 public:
-    using key_type            = typename CuckooTraits<This_t>::key_type;
-    using mapped_type           = typename CuckooTraits<This_t>::mapped_type;
+    using key_type       = typename CuckooTraits<This_t>::key_type;
+    using mapped_type    = typename CuckooTraits<This_t>::mapped_type;
     using iterator       = typename Base_t::iterator;
     using const_iterator = typename Base_t::const_iterator;
 

### 2017-03-17 include/cuckoo_independent2l.h

@@ -12,21 +12,22 @@ class CuckooIndependent2L
     : public CuckooTraits<CuckooIndependent2L<K,D,HF,Conf> >::Base_t
 {
 private:
-    using This_t         = CuckooIndependent2L<K,D,HF,Conf>;
-    using Base_t         = typename CuckooTraits<This_t>::Base_t;
-    using Bucket_t       = typename CuckooTraits<This_t>::Bucket_t;
-    using Hasher_t       = typename CuckooTraits<This_t>::Hasher_t;
-    using Hashed_t       = typename Hasher_t::Hashed_t;
-    using Ext            = typename Hasher_t::Extractor_t;
+    using This_t             = CuckooIndependent2L<K,D,HF,Conf>;
+    using Base_t             = typename CuckooTraits<This_t>::Base_t;
+    using Bucket_t           = typename CuckooTraits<This_t>::Bucket_t;
+    using Hasher_t           = typename CuckooTraits<This_t>::Hasher_t;
+    using Hashed_t           = typename Hasher_t::Hashed_t;
+    using Ext                = typename Hasher_t::Extractor_t;
 
     friend Base_t;
     friend iterator_incr<This_t>;
 
 public:
-    using key_type            = typename CuckooTraits<This_t>::key_type;
-    using mapped_type           = typename CuckooTraits<This_t>::mapped_type;
-    using iterator       = typename Base_t::iterator;
-    using const_iterator = typename Base_t::const_iterator;
+    using key_type           = typename CuckooTraits<This_t>::key_type;
+    using mapped_type        = typename CuckooTraits<This_t>::mapped_type;
+    using iterator           = typename Base_t::iterator;
+    using const_iterator     = typename Base_t::const_iterator;
+    using insert_return_type = typename Base_t::insert_return_type;
 
 private:
     using value_intern   = std::pair<key_type, mapped_type>;
@@ -110,12 +111,12 @@ private:
 public:
     // Specialized Funcitions (to keep per table counts) ***********************
 
-    std::pair<iterator, bool> insert(const key_type k, const mapped_type d)
+    inline insert_return_type insert(const key_type k, const mapped_type d)
     {
         return insert(std::make_pair(k,d));
     }
 
-    std::pair<iterator, bool> insert(const std::pair<key_type, mapped_type> t)
+    inline insert_return_type insert(const std::pair<key_type, mapped_type> t)
     {
         auto hash = hasher(t.first);
[…]

### 2017-03-17 include/cuckoo_simple.h

@@ -79,7 +79,7 @@ public:
                      : std::make_pair(0,nullptr);
     }
 
-    inline iterator begin() const
+    inline iterator begin()
     {
         auto temp = make_iterator(&table[0].elements[0]);
         if (! temp->first) temp++;

### 2017-03-17 include/hopscotch.h

@@ -74,8 +74,8 @@ public:
     inline size_t erase(const key_type& k)
     { return table.erase(k); }
 
-    inline iterator begin()  const { return table.begin();  }
-    inline iterator end()    const { return table.end();    }
+    inline iterator begin()  { return table.begin();  }
+    inline iterator end()    { return table.end();    }
     inline iterator cbegin() const { return table.cbegin(); }
     inline iterator cend()   const { return table.cend();   }
 

### 2017-03-17 include/prob_base.h

@@ -71,20 +71,26 @@ public:
     size_t                    erase (const key_type& k);
 
     // Easy use Accessors for std compliance ***********************************
-    inline iterator           begin () const
+    inline iterator           begin ()
     {
         auto temp = make_iterator(&table[0]);
         if (!temp->first) temp++;
         return temp;
     }
-    inline iterator           end   () const { return make_iterator(nullptr);  }
+    inline const_iterator     begin()  const
+    {
+        return static_cast<Specialized_t*>(this)->cbegin();
+    }
     inline const_iterator     cbegin() const
     {
         auto temp = make_citerator(&table[0]);
         if (!temp->first) temp++;
         return temp;
     }
+    inline iterator           end   () { return make_iterator(nullptr);  }
+    inline const_iterator     end   () const { return static_cast<Specialized_t*>(this)->cend(); }
     inline const_iterator     cend  () const { return make_citerator(nullptr); }
+
     mapped_type&              at    (const key_type& k);
     const mapped_type&        at    (const key_type& k) const;
     mapped_type&              operator[](const key_type& k);

### 2017-03-17 include/std_unordered.h

@@ -8,13 +8,14 @@ class STDProb
 {
 private:
     using Table_t  = std::unordered_map<K,D,HF>;
-    using iterator = typename Table_t::iterator;
 
     Table_t table;
 
 public:
     using key_type    = K;
     using mapped_type = D;
+    using iterator = typename Table_t::iterator;
+    using const_iterator = typename Table_t::const_iterator;
 
     STDProb(size_t cap, double = 1.1, size_t = 0) : table(cap) {}
 
@@ -38,8 +39,8 @@ public:
         return table.erase(k);
     }
 
-    inline iterator       begin()  const { return table.begin(); }
-    inline iterator       end()    const { return table.end();   }
+    inline iterator       begin()  { return table.begin(); }
+    inline iterator       end()    { return table.end();   }
     inline const_iterator cbegin() const { return table.cbegin(); }
     inline const_iterator cend()   const { return table.cend();   }
 

### 2017-04-10 paper/paper.tex
Laptop - Paper Stuff

\documentclass[a4paper,UKenglish]{lipics-v2016}
%This is a template for producing LIPIcs articles.
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
% for section-numbered lemmas etc., use "numberwithinsect"

\usepackage{microtype}%if unwanted, comment out or use option "draft"

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the recommended bibstyle

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Dynamic Space Efficient Cuckoo Table when doubling is too much -- a novel approach to growing!}
\titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{Tobias Maier}
\author[1]{Dr. Peter Sanders}
\affil[1]{Karlsruhe Institute of Technology, Karlsruhe, Deutschland\\
  \texttt{\{t.maier, sanders\}@kit.edu}}
%% \affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
%%   \texttt{access@dummycollege.org}}
\authorrunning{T. Maier and P. Sanders} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{Tobias Maier and Peter Sanders}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation".
\keywords{Dynamic data-structures, Open addressing, Closed hashing}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{hyperref}
%\usepackage{graphix}

\begin{document}

\maketitle

\begin{abstract}
A lot of previous work concentrates on optimizing the speed and space
efficiency of hash tables.  Additionally many practical
implementations exist.  Modern displacement strategies can fill a
given table to more than 95\% before the performance begins to decrease
significantly.  But once the threshold is met, copying and migrating
the whole table still seems to be the most efficient answer.

To achieve any kind of efficiency, each such migration must be
amortized by newly inserted elements.  Therefore, each migration has
to increase the capacity by a significant amount -- usually, this
means doubling the capacity.  Growing by a constant factor
$\gamma$ often leads to bad space efficiency because the newly grown table
is only filled to less than a factor of $\gamma^{-1}$.

The space efficiency is even worse during the migration, where both
the new and the old table coexist at the same time causing a massive
space overhead.  To solve both of these issues, we propose the a new
hash table architecture that supports dynamic growing while keeping a
strict size constraint $\alpha \cdot n$ throughout its lifetime.  Even
under these conditions, we can guarantee constant lookup times, and
amortized insertion times, independent of $n$.
 \end{abstract}

%% \section{Typesetting instructions -- please read carefully}
%% Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume.
%% \begin{itemize}
%% \item Use pdflatex and an up-to-date LaTeX system.
%% \item Use further LaTeX packages only if required. Avoid usage of packages like \verb+enumitem+, \verb+enumerate+, \verb+cleverref+. Keep it simple, i.e. use as few additional packages as possible.
%% \item Add custom made macros carefully and only those which are needed in the article (i.e., do not simply add your convolute of macros collected over the years).
%% \item Do not use a different main font. For example, the usage of the \verb+times+-package is forbidden.
%% \item Provide full author names (especially with regard to the first name) in the \verb+\author+ macro and in the \verb+\Copyright+ macro.
%% \item Fill out the \verb+\subjclass+ and \verb+\keywords+ macros. For the \verb+\subjclass+, please refer to the ACM classification at \url{http://www.acm.org/about/class/ccs98-html}.
%% \item Take care of suitable linebreaks and pagebreaks. No overfull \verb+\hboxes+ should occur in the warnings log.
%% \item Provide suitable graphics of at least 300dpi (preferrably in pdf format).
%% \item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsection*+, \verb+\paragraph+, \verb+\subparagraph*+, ... ``Self-made'' sectioning commands (for example, \verb+\noindent{\bf My+ \verb+subparagraph.}+ will be removed and replaced by standard LIPIcs style sectioning commands.
%% \item Do not alter the spacing of the  \verb+lipics-v2016.cls+ style file. Such modifications will be removed.
%% \item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
%% \item Remove all comments, especially avoid commenting large text blocks and using \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions.
%% \item Keep the standard style (\verb+plainurl+) for the bibliography as provided by the\linebreak \verb+lipics-v2016.cls+ style file.
%% \item Use BibTex and provide exactly one BibTex file for your article. The BibTex file should contain only entries that are referenced in the article. Please make sure that there are no errors and warnings with the referenced BibTex entries.
%% \item Use a spellchecker to get rid of typos.
%% \item A manual for the LIPIcs style is available at \url{http://drops.dagstuhl.de/styles/lipics-v2016/lipics-v2016-authors/lipics-v2016-manual.pdf}.
%% \end{itemize}

\section{Introduction}
Hash tables are some of the most frequently used data-structures. They
belong into every programmers basic toolbox.  As such having
interchangeable implementations which perform well under different
conditions is important and allows programmers
easily adapt known solutions to new circumstances.

One aspect which a lot of research focuses on is space efficiency.
Modern space efficient hash tables work well up to loads of 98\% and
more. But for this to matter, programmers have to know similarly tight
bounds to the absolute number of elements inserted into the table.
This is typically not the case, one of the most ubiquitous use cases
for hash tables is to unify data according to some key member.  To
guarantee good performance, we have to overestimate the necessary
capacity whenever the exact number of unique keys is not known a
priori.  Dynamic data-structures are necessary to alleviate this
problem.  In circumstances where space efficiency is important, space
efficiently growing hash tables are necessary.

Many libraries -- even ones that implement space efficient hash tables
-- already offer dynamic growing.  The problem with these
implementations is that they either lose their space efficiency or
their overall performance, once the table grows above its original
capacity.  Usually growing is implemented by either creating
additional hash tables -- decreasing the performance especially for
look ups, or by migrating all elements to a new table -- losing the
space efficiency by multiplying the original size.

To avoid both of these pitfalls we propose a variant of bucket cuckoo
hashing (TODO REFERENCE).  Bucket cuckoo hashing is a technique where
each element can be stored in one of several associated constant sized
buckets.  When all of them are full we move elements, to create new
space.  We split one bucket cuckoo table into multiple subtables each
of which can grow independent from all others.  Because the buckets
associated with one element are spread over the different subtables,
growing one subtable alleviates pressure from all other tables.

Doubling the size of one subtable increases the overall size by a
small factor, while only moving a small number of elements -- making
it possible, to amortize these small size changes. The introduced size
and occupancy imbalance between subtables is alleviated using
displacement techniques common to cuckoo hashing. Using these
techniques, makes our table work and grow efficiently with fill
degrees around 97\% (TODO READ APPROPRIATE VALUE) .  Being
approximately \ldots{} \% faster than common hash table techniques under the
same circumstances.

\section{Motivation}
Space efficiency is a major factor for the development of new hashing
techniques.  The goal has long been to make hash tables work better in
high load scenarios.  Modern hashing techniques like Hopscotch, Robin
Hood, and Cuckoo Hashing can easily fill hash tables to more than 90\%
and up to 98\% while still achieving constant average running times on
most of their operations.  But since these techniques depend on
statically sized tables this only makes sense, if the number of unique
elements can be estimated tightly.

To visualize this assume the following scenario.  During the operation
of a word count benchmark, we know an upper bound $c$ to the number of
unique words. Therefore, we construct a table large enough to hold
these elements ($0.9^{-1}\cdot c$).  It is entirely possible that the
specific word count instance only contains $0.7\cdot c$ unique words.
The resulting table is therefore only filled 63\%.  It is clear that
better static sized tables cannot solve this issue.

Depending on the underlying problem it is impossible to tightly
estimate the final number of elements.  Dynamic growing can be a
solution even in these instances.  It is completely independent from
the problem's domain and thus also works in cases where estimating the
number of elements can be hard.  But to achieve space efficiency in
scenarios where the final size is not known the hash table has to grow
closely with the actual number of elements.  This is not efficiently
possible with any of the growing techniques used currently.

\section{Related Work}
\begin{itemize}
\item Some other engineering papers

\item Some Space efficient Stuff  (theoretical)

\item Some Space efficient Stuff  (practical)
\end{itemize}

\section{Preliminaries}
A hash table is a data-structure, which stores key-value-pairs
(\(\langle key, data \rangle\)) and offers the following functions:
\verb~insert~ stores a given key-value pair, \verb~find~ given a key returns if
said key was stored, and if so it returns the stored value, and
\verb~erase~ which deletes a previously inserted element (if present).

An alternative model which is sometimes used in literature considers
arbitrary elements instead of key value pairs.  In that model, the key
is extracted using an extractor function.  All techniques presented in
this paper also work in the extractor model without any adaptation.

Throughout this paper, we use $n$ to denote the number of elements and
$m$ the number of cells ($m > n$) in any given hash table.  We define
the load factor as $load = n/m$.  Tables can usually only work
efficiently up to a certain load factor.  Above that, operations get
slower and/or have a possibility to fail.
When implementing a hash table one has to decide between storing
elements directly in the hash table -- \emph{Closed Hashing} -- or storing
pointers to elements -- \emph{Open Hashing}. This has an immediate impact
on the necessary memory.

\begin{align*}
  mem_{closed}(n,m) &\geq^* m\cdot size(element)\\
  mem_{open}  (n,m) &\geq^* m\cdot size(pointer) + n\cdot size(element)\\
\end{align*}

For large elements (\(\gg size(pointer)\)), one can simply use open
hashing, to reduce the relevant memory factor.  Therefore, we restrict
ourselves to the common case of elements whose size is comparable to
that of a pointer.  For our experiments we use 128bit elements (64bit
keys and 64bit values).  In these instances closed addressing
introduces a significant memory overhead (at least \(\times1.5\)),
therefore, we only consider closed hash tables for the purpose of this
paper. For closed hash tables, the memory efficiency is directly
dependent on the element density.

\subsection{$\alpha$-Space Efficient Hash Tables}
We call a hashing technique \(\alpha\)-space efficient when it can
work effectively using at most $\alpha \cdot size_{\min} + O(1)$
memory (TODO DEFINE WORKING EFFICIENTLY FURTHER i.e. constant average
running time or just look at experiments or inserting the last 10\%
elements is less than inserting all others). In this case we define
working efficiently as having average insertion times close to
$O(\frac{m}{m-n})$ (average number of probes to hit an empty cell).

\[\alpha \cdot size_min = \alpha\cdot n\cdot size(element)\]

In many closed hashing techniques -- i.e. linear probing, cuckoo
hashing \ldots{} -- cells are the same size as elements. Therefore, being
$\alpha$ space efficient is the same as operating with a load factor
of $\alpha^{-1} = n/m$ ($\alpha\cdot n \cdot size(element) = m\cdot
size(element)$).  This is not necessarily the case for every kind of
hashing technique.  Some techniques like hopscotch hashing use
additional per cell information.  This would also be the case if we
store some per bucket data like the number of stored elements in our
bucket cuckoo hash table.

\subsubsection*{$\alpha$-Space Efficiency for Dynamic Tables}
The definition of a space efficient hashing technique given in (TODO
reference) the previous section is specifically targeted for
statically sized hash tables.  As detailed in Motivation (TODO
REFERENCE) our goal is to construct a hash table which can efficiently
(and tightly) grow with the number of inserted elements. Therefore, we
want to define the notion of an \(\alpha\)-space efficient dynamic
table.

We call a hash table implementation dynamically \(\alpha\)-space
efficient if an instantiated table can grow arbitrarily large over its
original capacity, while remaining smaller than \(\alpha\cdot
mem_{necessary}\) at all times.  \[mem_{curr} \leq \alpha\cdot
max(n_{curr})\cdot size(element)\]

One point of discussion while defining dynamic space efficiency is the
space consumption during hash table operations.  During a table
migration both the source (\(t_{source}\)) and the target
(\(t_{target}\)) table are allocated, thus using \(mem =
(m_{source}+m_{target})\cdot size(cell)\) space.  This is especially
prohibiting if all elements are held in one table, then at least
\(2\cdot m\cdot size(element)\) is necessary for each migration.

Our definition of dynamic space efficiency does not enforce size
reduction due to the deletion of elements, but it enforces that the
memory used by deleted elements must be reused (no tombstones).

\paragraph{Trivial Examples}
One simple example of this is a linear probing table, which grows by a
factor of two, whenever 50\% of the table is filled. Here the table is
at least 25\% filled (immediately after growing), therefore, the table
is 6-space efficient (during the grow operation there are \((4+2)+n\)
cells). Another example would be a cuckoo table which doubles in size
whenever it surpasses 95\%, fill rate. This would be a
$0.95^{-1}+0.45^{-1}\approx 3.2$-space efficient table.

\subsection{Cuckoo Hashing}
Cuckoo hashing is a technique to resolve hash conflicts in an closed hashing hash table. Its main draw is that it guarantees constant
lookup times even in densely filled tables. The article TODO by TODO
(probably mitzenmacher with or without dietzfelbinger) et al. gives a
good overview over many cuckoo hashing variants.  The distinguishing
technique of cuckoo hashing is, that $H$ hash functions (\(h_1, ... ,
h_H\)) are used to find $H$ independent possible positions. Each
element is stored in one of its positions.  Even if all possible
positions are occupied one can often move elements to create space for the
current element. We call this displacing elements.

Bucket cuckoo hashing is a variant, where the cells of the hash table
are grouped into buckets of size $B$ ($m/B$ buckets).  Each element
appointed to one bucket can be stored in any of the bucket's cells .
Using buckets one can increase the number of displacement
opportunities drastically.

\textbf{Lookup:} Find and Remove operations have a guaranteed constant
running time. Independent from the tables density, there are $H$
possible buckets -- $H\cdot B$ cells -- that have to be searched to
find an element.

\textbf{Insert:} Each element is hashed to $H$ buckets.  If at least
one of those buckets has space left, then we store the element in the
bucket, that has the most space.  But when all buckets are full, then we
have to move elements around the table, such that a new space becomes
available.

To visualize the problem of displacing elements, one can think of the
directed graph implicitly defined by the hash table.  Each bucket
defines one node and each element defines pairwise connections between
the bucket it is stored in to its $H-1$ alternate buckets.  To insert
an element into the hash table we have to find a path from one of its
associated buckets, to a bucket, that has space remaining.  Then we
move elements along this path, to open a space on the initial bucket.
The two common techniques to find such paths are \emph{random walks} and
\emph{breadth first searches}.

\subsubsection*{Some Performance Bounds}
\begin{itemize}
\item bla how dense can we get
\end{itemize}

When randomly probing the table we expect to find a free cell after $\frac{1}{1-load factor}$

\section{DySECT Space Efficient Growing, by Splitting the Table}
Our goal is to build an efficiently growing data structure, that
remains space efficient at all times.  A commonly used growing
technique is to double the previous size by migrating all elements
into a table which has two times the original size.  Of course this is
not memory efficient.  A natural thought which occurs is to double
only part of the data structure.  This idea together with the
balancing ability of multiple hash functions is what will achieve the
described functionality.

\subsection{Description}
We use a two level approach (shown in TODO FIGREFRENCE) to allow
differentiated growing between table parts.  The first level consists
of a table which stores pointers to $T$ second level tables. The
second level tables consist of buckets that store elements.

Comparable to classic cuckoo hashing, each element has $H$ associated
buckets ($H$ number of hash functions), these can be in the same, or
in different sub tables.  Each inserted element lies in one of its
associated buckets.  To find a bucket associated with an
element $e$, we compute \(e\)s hash value using the appropriated hash function
$h_i(e)$. We split that hash value into two parts, then we use the
first part to chose a second level table, and the second part to
choose one of the contained buckets.

To understand the performance of this hash table, it is important to
realize that when the hash table is used regularly, the first level
table will always remain cached. This is important, because otherwise lookups
would cause unnecessary cache misses that would not have happened in a
single level hash table.  Because of this we expect the average number
of cache misses during hash table operations to be similar to those of
a normal cuckoo hash table.

\subsection{Growing}
When the table contains enough elements, that the memory constraint
can be kept, we migrate a subtable.  We migrate subtables in order
from the first to the last, therefore, no subtable can be more than
twice as large as any other.

Assume that all subtables have $s=m/tl$ cells. We can grow the first
subtable once $\alpha\cdot n > m+2s$.  Doubling the size of a second
level table increases the global number of cells from $m_{old} =
tl\cdot s$ to $m_{new} = m_{old}+s = (tl+1)\cdot s$ (factor:
$\frac{tl+1}{tl}$).  Note that all subsequent grows migrate one of the
smaller tables, until all tables have the same size again.  Therefore,
each grow until then increase the capacity by the same absolute amount
(smaller factor).

From a theoretical point of view, the cost of growing the subtable is
amortized by element insertions.  There are at least $\alpha^{-1}
\cdot s = \Omega(s)$ insertions between two migrations.  One migration
takes $\Theta(s)$ time. Alternatively abstract point of view, after
inserting $\alpha^{-1} \cdot s$ elements the table grows enough, to
raise the capacity about the same $\alpha^{-1} \cdot s$ effective
cells (actual cells $\times$ fill factor).

From a practical view, the migration is efficient because it accesses
cells in a linear fashion making it really cache efficient.  Even in
the target table cells are accessed linearly, because of the way we
assign elements to buckets there are no displacements necessary.  The
elements from each original bucket are split into two buckets of the
target table.  Therefore, no bucket of the target table can have more
elements than its respective original bucket.

In the implicit graph model of the cuckoo table (TODOsee section
cuckoo hashing), growing a subtable is equivalent to
splitting each node that represents a bucket within that table. The
edges (elements) in the implicitly defined subgraph are not doubled,
therefore, the resulting subgraph becomes more sparse, making it
easier to insert elements.

\subsection{Shrinking}
In many use cases, shrinking is not necessary.  It worsens performance
by taking time for the migration and making the remaining table more
dense.  Furthermore, it is unclear that the freed memory is actually
necessary for other parts of the application.  If that memory is
necessary, and shrinking is required, for example, when removed
elements are reinserted into another data structure and global memory
bounds are necessary (other data structures grow while the hash table
shrinks).

Shrinking can work similarly to growing. We replace a second level
table with a smaller one by migrating elements from one to the
other.  During this migration we join elements from two buckets into
one. Therefore, it is possible for a bucket to overfill.  We reinsert
these elements at the end of the migration.  Obviously, this can only
apply to at most half the migrated elements.

When triggering the size reduction, one has to make sure that the
migration is amortized. Therefore, a grow operation cannot immediately
follow a shrink operation.  When shrinking is enabled we propose to
shrink one subtable when $\alpha*n < m-s'$ elements ($s'$ size of a
large table, $m_{new} = m_{old} -s'/2$).

Notice that the memory requirement actually increases during the
operation itself.
\subsection{Hashing Elements to Buckets}
\subsubsection*{Compute the Bucket from a Hashed Key}
From a hashed key, we have to compute the subtable and the bucket
within that subtable.  To make this efficient we use powers of two for
the number of subtables ($T = 2^t$), as well as the individual
subtable sizes ($s = 2^*$).  Since the number of subtables remains
constant we can use the first $t$ bits from the hashed key, to find the
appropriate subtable.  From the remaining bits we compute the offset
into this subtable using a bitmask ($\texttt{AND}~s-1 \leftrightarrow
\mod s$).

\subsubsection*{Reducing Number Computed Hash Functions}
Evaluating hash functions is expensive, therefore, reducing the number
of hash functions computed per operation increases the efficiency of
the table.  The hash function we use computes 64bit hash values
(i.e. xxHash TODO).  We split these 64bit value into two 32bit hash
values. All common bucket hash table sizes can be addressed using 32
bits. They can address up to $2^{32}$ buckets which achieves a
capacity for \(2^{35} \approx 34\) billion elements (bucket size 8),
at that point a hash table consumes 512GiB memory.

When $H > 2$ we can use \emph{double hashing} to further reduce the number
of computed hash functions. Double hashing creates an arbitrary amount
of hash values using only two original hash functions $h'$ and
$h''$.  The additional values are linear combinations computed
from the original two values, $h_i(key) = h'(key) + i\cdot
h''(key)$.

Combining both of these techniques, we can reduce the number of
computed hash functions to one 64bit hash function.  This is
especially important during large displacements, where each
encountered element has to be rehashed to find its alternative
buckets.

\subsection{Difficulties for the Analysis of our table structure}
There are three factors, that impact the performance of our dynamic
table compared to other cuckoo table variants and to other hashing solutions in
general \textbf{inhomogeneous table resolution}, \textbf{element imbalance}, and
\textbf{population density}. All of these factors influence the maximum load
density and the running times in different ways.

\subsubsection*{Imbalance through Inhomogeneous Table Resolution}
\label{sec:inhom_res}
By growing individual second level tables we introduce a size imbalance
between subtables.  Large tables contain more buckets, but the number
of elements hashed to the table itself is not dependent on its size,
therefore, it is difficult to spread elements evenly among buckets.
Uneven bucket fill ratios can lead to longer insertion times.

If there are $n$ elements in a table with \emph{tl} second level tables,
$j$ of which have size $2s$ the others have size $s$. If elements are
spread equally among buckets then all small tables have around
$n/(tl+j)$ elements, and the bigger tables have $2n/(tl+j)$ elements.
For each table there are about $kn/tl$ elements associated to one of
its buckets.  This shows that having more hash functions can lead to a
better balance.

For two hash functions ($k=2$) and only one grown table ($j=1$) this
means that $\approx 2n/(tl+1)$ should be stored in the first table.
These are nearly all elements associated to a bucket in the first
table ($\approx 2n/tl$). So to distribute elements evenly nearly all
elements would have to be stored there.

\subsubsection*{Imbalance through Size Changes}
In addition to the problem of inhomogenous tables, there is an
inherent balancing problem introduced by resizing subtables. It is
clear, that a newly grown table is not filled as densely as other
tables.  Since we double the table size, grown tables can only be
filled to about 50\%.

Assume the global table is filled close to 100\% when the first table begins
to grow.  now there is the capacity for $s$ new elements, but this
capacity is only in the first table, elements that are not hashed to
the first table, automatically trigger displacements leading to slow
insertions.

Notice that repeated operations help to equalize this imbalance,
because elements are more likely inserted into the less dense areas,
and more likely to be deleted from more dense areas.

\subsubsection*{Population Density}
Measuring and comparing the performance of space efficient growing
tables to their statically sized counterparts is difficult.  Inserting
elements into a densely filled table takes longer, than into an empty
one.  Therefore, a table that is always densely filled will naturally
be slower than one that only fills up towards the end of the
execution -- even when ignoring eventual growing costs.

\section{Experiments}
As described in section (TODO REFERENCE DIFFICULTIES\ldots{}) there are
many factors that impact hash table performance.  Many of these
factors are inherent to the combination of growing and space
efficiency.  To show that our ideas work from a practical point of
view we use a wide portfolio of both micro-benchmarks and practical
experiments.

\subsection{Comparison Implementations}
To put the performance of our solution into perspective, we
implemented and tested several other options for space efficient
hashing.  We use our own implementations, since most hash tables found
online do not support strict space-efficiency.  In Appendix TODO we
show that our implementations are at least competitive with
implementations found .

\textbf{Linear Probing} is the most common hashing technique outside of
hashing with chaining.  It is commonly accepted to be inefficient for
higher load factors because search distances can become large.  In
practice, this fact is counteracted by its cache-efficiency making
operations fast in spite of long probing distances.  Linear probing is
especially slow in \verb~find~ and \verb~erase~ operations, that often have
guaranteed constant running time in other implementation.

To grow our linear probing table, we use table migrations.  This can
be made space efficient, by increasing the capacity in place.  This is
possible due to over allocation (MAYBE SOME SOURCE/APPENDIX/WRITE
SHORT PAPER) and a simple in place migration algorithm (ALSO
APPENDIX/SHORT PAPER).  For a table that was initialized to hold $n'$
elements with a space-efficiency $\alpha$ we trigger growing once the
number of elements is $n = \frac{(1+\alpha)}{2}n'$. We then increase
the capacity $m$ to $\alpha\cdot n$ to remain strictly size-efficient.
Our implementation of the migration is highly cache efficient, but
all elements have to be moved.  This makes repeated migrations inefficient
especially for small $\alpha$.

\textbf{Hopscotch Hashing} is a technique similar to linear probing where
elements are moved when they are too far from their original hash
position.  Additionally, hopscotch hashing introduces an acceleration
data structure which uses bitmaps to store neighborhood data. This
data improves find and erase performance.  Many people would expect
the acceleration data to be the main draw to hopscotch hashing, but it
is also its greatest demise in the space-efficient setting.  The
bitmaps introduce a significant memory overhead of neighborhood-size
bits per cell.  Our experiments show that for any reasonable fill
factor neighborhoods have to be 64bits and larger, this introduces a
size factor of $\times1.5$ (128bit elements).  We still include
measurements with hopscotch hashing -- without considering space for
the neighborhood data -- to show its overall performance tendencies.
Growing hopscotch tables is done similar to the technique described
under linear probing.

\textbf{Robin Hood Hashing} is another technique based on linear probing. The
goal of this technique is to improve find performance by minimizing
the maximum/average displacement distance.  The drawback of this technique is
that insertions have to compute the hash of each element that is
skipped by the insertion.

\textbf{Classic Bucket Cuckoo Hashing} is the technique we adopted to design
our hashing technique.  It has clear advantages for space efficient
tables.  Like our variant it offers constant lookup and deletion
times.  In contrast to our variant, it does not suffer from any
imbalances caused either by table resolution or by local growing
operations.  Growing works similar to the linear probing table above.

\subsection{Hardware}
We test our implementation on two very different systems, first a
consumer level desktop computer \ldots{}

\subsection{Tests with a Static Table Size}
Our hash table is designed for space efficient growing, therefore, its
main advantages show in dynamic workloads.  We still performed some
tests with a static table size.  With these measurements, we can
highlight different points without the influence of dynamically growing.

The following test was performed, by initializing a table with
$m\approx24M$~cells (non-growing). Then elements are inserted, until there
is a critical amount of failing insertions ($>100$).  At different
stages, we measure the speed of new insertions
(Figure~\ref{fig:eps_insert}), and finds (Figure~\ref{fig:eps_find}).
Performance is measured by performing each operation 1000 times, finds
are measured using either random elements within the table (successful), or by
searching random elements from key space (unsuccessful).

As one might expect, the performance, of insert operations depends
highly on the fill degree of the table.  Therefore, we show it
normalized with $\frac{1}{\varepsilon}$, which is the expected number
of fully random probes, to find a free cell. We see that -- up to a
certain point -- for all tables the insertion time behaves
proportional to $\frac{1}{\varepsilon}$, then it begins to decrease
uncontrolled.  The plot shows, that our table remains stable for
longer than linear probing variants, but not as long, as for classic
cuckoo hashing.  Hopscotch hashing fails earlier than all other
tables, because its bitmaps become to small, to store the neighborhood
information.

DySECT fails earlier than cuckoo hashing.  The cause for this is
twofold. First, our architecture ($T$ subtables with power of two
sizes) does not allow exactly $24M$ cells, therefore, our table is
slightly more full. Secondly, to get close to $24M$ cells, we have
$110$ large subtables (twice the size of the others).  This introduces
inhomogeneous table resolution (see \ref{sec:inhom_res}), which
decreases our performance compared to cuckoo hashing.  This test shows
the influence of inhomogeneous table resolution without the influence
of size changes, and the connected table imbalance.

Figure~\ref{fig:eps_find} shows that linear probing variants have bad
performance on find operations on highly filled tables.  It also shows
that unsuccessful find operations are often very slow when using
linear probing.  This is where the acceleration data-structure used by
hopscotch tables can significantly improve performance.  Both of these
effects, do not affect cuckoo variants like DySECT, that have
guaranteed constant running times for all find operations --
independant of the fill degree.

\subsubsection*{Constant Sized Table Under Varying Fill Degree}
\begin{figure}
  \includegraphics[width=\textwidth]{../eval/plots/eps_insert.pdf}
  \caption{Insert timings at different fill degrees into a non-growing table.}
  \label{fig:eps_insert}
\end{figure}
\begin{figure}
  \includegraphics[width=0.48\textwidth]{../eval/plots/eps_find_s.pdf}
  \includegraphics[width=0.48\textwidth]{../eval/plots/eps_find_u.pdf}
  \caption{Timings for find operations with different fill degrees (\emph{left:}successful, \emph{right:}unsuccessful).}
  \label{fig:eps_find}
\end{figure}

\subsubsection*{Constructing a Table with Growing}

\begin{figure}
  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
  \caption{Average time per insert operation during the construction of a dynamic table.}
  \label{fig:ti_insert}
\end{figure}
\begin{figure}
  \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_s.pdf}
  \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_u.pdf}
  \caption{Timings for find operations on the grown table (from Figure~\ref{fig:ti_insert}; \emph{left:}successful, \emph{right:}unsuccessful).}
  \label{fig:ti_find}
\end{figure}

%%
%% Bibliography
%%

%% Either use bibtex (recommended),

\bibliography{lipics-v2016-sample-article}

%% .. or use the thebibliography environment explicitely

\end{document}

### 2017-04-11 include/prob_hops.h
Laptop - hipprob now uses buckets -> bigger neighborhood

@@ -93,7 +93,8 @@ private:
     using This_t = HopsProb<K,D,HF,Conf>;
     using Base_t = typename ProbTraits<This_t>::Base_t;
 
-    static constexpr size_t nh_size = Conf::NeighborSize;
+    static constexpr size_t nh_size = (Conf::NeighborSize <= 64) ?
+                                       Conf::NeighborSize : 64;
     using AugData_t = AugmentData<nh_size>;
 
     friend Base_t;
@@ -326,7 +327,9 @@ private:
     using Base_t = typename ProbTraits<This_t>::Base_t;
 
     static constexpr size_t nh_size = Conf::NeighborSize;
-    using AugData_t = AugmentData<nh_size>;
+    static constexpr size_t bitset_size = (nh_size < 64) ? nh_size : 64;
+    static constexpr size_t bucket_size = (nh_size < 64) ? 1 : nh_size/64;
+    using AugData_t = AugmentData<bitset_size>;
 
     friend Base_t;
 
@@ -349,9 +352,11 @@ public:
         table = std::unique_ptr<value_intern[]>(temp);
 
         capacity = (cap) ? cap*alpha : 2048*alpha;
+        capacity = capacity/bucket_size;
+        capacity = capacity*bucket_size;
         thresh   = (cap) ? cap*beta  : 2048*beta;
 
-        factor = double(capacity-nh_size)/double(1ull << 32);
+        factor = double(capacity/bucket_size-bitset_size)/double(1ull << 32);
 
         std::fill(table.get(), table.get()+capacity, value_intern());
         nh_data.clear_init(capacity);
@@ -394,15 +399,18 @@ public:
         auto   aug  = nh_data.getAcc(ind);
         size_t bits = aug.getNHood();
 
-        for (size_t i = ind; bits; ++i, bits>>=1)
+        for (size_t i = ind; bits; i+=bucket_size, bits>>=1)
         {
             if (!(bits & 1)) continue;
             else
             {
-                auto temp = table[i];
-                if ( temp.first == t.first )
+                for (size_t ti = 0; ti<bucket_size; ++ti)
                 {
-                    return std::make_pair(make_iterator(&table[i]), false);
+                    auto temp = table[i+ti];
+                    if ( temp.first == t.first )
+                    {
+                        return std::make_pair(make_iterator(&table[i]), false);
+                    }
                 }
             }
         }
@@ -420,7 +428,7 @@ public:
                     if (!successful) break;
                 }
[…]

### 2017-04-11 source/selection.h

@@ -269,6 +269,8 @@ struct Chooser
         {
         case 64:
             return executeN<Functor,64>(c, std::forward<Types>(param)...);
+        case 128:
+            return executeN<Functor,128>(c, std::forward<Types>(param)...);
         // case 8:
         //     return executeN<Functor, 8>(c, std::forward<Types>(param)...);
         // case 16:

### 2017-04-11 include/prob_hops.h
Laptop - small fix to hipprob erase

@@ -493,7 +493,7 @@ public:
                     for (size_t tti = 0; tti < bucket_size; ++tti)
                     {
                         auto ttk = table[i+tti].first;
-                        if (h(k) == ind) return 1;
+                        if (h(ttk) == ind) return 1;
                     }
                     nh_data.getAcc(ind).unset((i-ind)/bucket_size);
                     return 1;

### 2017-04-11 paper/paper.tex
Laptop - Paper Stuff

@@
 E.g., cite as "F.1.1 Models of Computation".
-\keywords{Dynamic data-structures, Open addressing, Closed hashing}% mandatory: Please provide 1-5 keywords
+\keywords{Dynamic data-structures, Open addressing, Closed hashing, Cuckoo hashing, Dense hash tables}% mandatory: Please provide 1-5 keywords
 % Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 efficiency of hash tables.  Additionally many practical
-implementations exist.  Modern displacement strategies can fill a
-given table to more than 95\% before the performance begins to decrease
+implementations exist.  Modern open addressing techniques can fill a
+given table to more than 95\% before the performance deteriorates
 significantly.  But once the threshold is met, copying and migrating
@@
 to increase the capacity by a significant amount -- usually, this
-means doubling the capacity.  Growing by a constant factor
+means doubling the capacity.  Growing by large factor
 $\gamma$ often leads to bad space efficiency because the newly grown table
@@
 
-The space efficiency is even worse during the migration, where both
-the new and the old table coexist at the same time causing a massive
-space overhead.  To solve both of these issues, we propose the a new
-hash table architecture that supports dynamic growing while keeping a
-strict size constraint $\alpha \cdot n$ throughout its lifetime.  Even
-under these conditions, we can guarantee constant lookup times, and
-amortized insertion times, independent of $n$.
+The space efficiency is even worse during the migration, where
+oftentimes both the new and the old table coexist at the same time
+causing a massive space overhead.  To solve both of these issues, we
+propose the a new hash table architecture that supports dynamic
+growing while keeping a strict size constraint $\alpha \cdot n$
+throughout its lifetime.  Even under these conditions, we can
+guarantee constant lookup times, and amortized insertion times,
+independent of $n$.
  \end{abstract}
@@
 A hash table is a data-structure, which stores key-value-pairs
-(\(\langle key, data \rangle\)) and offers the following functions:
+($\langle key, data \rangle$) and offers the following functions:
 \verb~insert~ stores a given key-value pair, \verb~find~ given a key returns if
@@
 
-For large elements (\(\gg size(pointer)\)), one can simply use open
+For large elements ($\gg size(pointer)$), one can simply use open
 hashing, to reduce the relevant memory factor.  Therefore, we restrict
@@
 keys and 64bit values).  In these instances closed addressing
-introduces a significant memory overhead (at least \(\times1.5\)),
+introduces a significant memory overhead (at least $\times1.5$),
 therefore, we only consider closed hash tables for the purpose of this
@@
 \subsection{$\alpha$-Space Efficient Hash Tables}
-We call a hashing technique \(\alpha\)-space efficient when it can
+We call a hashing technique $\alpha$-space efficient when it can
 work effectively using at most $\alpha \cdot size_{\min} + O(1)$
@@
 (and tightly) grow with the number of inserted elements. Therefore, we
-want to define the notion of an \(\alpha\)-space efficient dynamic
+want to define the notion of an $\alpha$-space efficient dynamic
 table.
 
-We call a hash table implementation dynamically \(\alpha\)-space
+We call a hash table implementation dynamically $\alpha$-space
 efficient if an instantiated table can grow arbitrarily large over its
-original capacity, while remaining smaller than \(\alpha\cdot
-mem_{necessary}\) at all times.  \[mem_{curr} \leq \alpha\cdot
-max(n_{curr})\cdot size(element)\]
+original capacity, while remaining smaller than $\alpha\cdot
+mem_{necessary}$ at all times.
+\[mem_{curr} \leq \alpha\cdot max(n_{curr})\cdot size(element)\]
 
@@
 space consumption during hash table operations.  During a table
-migration both the source (\(t_{source}\)) and the target
-(\(t_{target}\)) table are allocated, thus using \(mem =
-(m_{source}+m_{target})\cdot size(cell)\) space.  This is especially
+migration both the source ($t_{source}$) and the target
+($t_{target}$) table are allocated, thus using $mem =
+(m_{source}+m_{target})\cdot size(cell)$ space.  This is especially
 prohibiting if all elements are held in one table, then at least
-\(2\cdot m\cdot size(element)\) is necessary for each migration.
+$2\cdot m\cdot size(element)$ is necessary for each migration.
 
@@
 
-\paragraph{Trivial Examples}
+\paragraph*{Trivial Examples}
 One simple example of this is a linear probing table, which grows by a
@@
 at least 25\% filled (immediately after growing), therefore, the table
-is 6-space efficient (during the grow operation there are \((4+2)+n\)
+is 6-space efficient (during the grow operation there are $(4+2)+n$
 cells). Another example would be a cuckoo table which doubles in size
@@
 \subsection{Cuckoo Hashing}
-Cuckoo hashing is a technique to resolve hash conflicts in an closed hashing hash table. Its main draw is that it guarantees constant
+Cuckoo hashing is a technique to resolve hash conflicts in an closed
+hashing hash table. Its main draw is that it guarantees constant
 lookup times even in densely filled tables. The article TODO by TODO
@@
 good overview over many cuckoo hashing variants.  The distinguishing
-technique of cuckoo hashing is, that $H$ hash functions (\(h_1, ... ,
-h_H\)) are used to find $H$ independent possible positions. Each
+technique of cuckoo hashing is, that $H$ hash functions ($h_1, ... ,
+h_H$) are used to find $H$ independent possible positions. Each
 element is stored in one of its positions.  Even if all possible
-positions are occupied one can often move elements to create space for the
-current element. We call this displacing elements.
+positions are occupied one can often move elements to create space for
+the current element. We call this displacing elements.
 
@@
 remains space efficient at all times.  A commonly used growing
-technique is to double the previous size by migrating all elements
-into a table which has two times the original size.  Of course this is
+technique is to increase the size by migrating all elements
+into a larger new table.  Of course this is
 not memory efficient.  A natural thought which occurs is to double
@@
 \subsection{Description}
-We use a two level approach (shown in TODO FIGREFRENCE) to allow
-differentiated growing between table parts.  The first level consists
-of a table which stores pointers to $T$ second level tables. The
-second level tables consist of buckets that store elements.
+We use multiple ($T$) subtables (shown in TODO FIGREFRENCE) to allow
+differentiated growing between table parts.  The subtables consist of
+buckets that store elements.
 
@@
 buckets ($H$ number of hash functions), these can be in the same, or
-in different sub tables.  Each inserted element lies in one of its
+in different subtables.  Each inserted element lies in one of its
 associated buckets.  To find a bucket associated with an
-element $e$, we compute \(e\)s hash value using the appropriated hash function
+element $e$, we compute $e$s hash value using the appropriated hash function
 $h_i(e)$. We split that hash value into two parts, then we use the
-first part to chose a second level table, and the second part to
+first part to chose a subtable, and the second part to
 choose one of the contained buckets.
 
-To understand the performance of this hash table, it is important to
-realize that when the hash table is used regularly, the first level
-table will always remain cached. This is important, because otherwise lookups
-would cause unnecessary cache misses that would not have happened in a
-single level hash table.  Because of this we expect the average number
-of cache misses during hash table operations to be similar to those of
-a normal cuckoo hash table.
+%% To understand the performance of this hash table, it is important to
+%% realize that when the hash table is used regularly, the first level
+%% table will always remain cached. This is important, because otherwise lookups
+%% would cause unnecessary cache misses that would not have happened in a
+%% single level hash table.  Because of this we expect the average number
+%% of cache misses during hash table operations to be similar to those of
+%% a normal cuckoo hash table.
 
@@
 
-Assume that all subtables have $s=m/tl$ cells. We can grow the first
-subtable once $\alpha\cdot n > m+2s$.  Doubling the size of a second
-level table increases the global number of cells from $m_{old} =
-tl\cdot s$ to $m_{new} = m_{old}+s = (tl+1)\cdot s$ (factor:
-$\frac{tl+1}{tl}$).  Note that all subsequent grows migrate one of the
-smaller tables, until all tables have the same size again.  Therefore,
-each grow until then increase the capacity by the same absolute amount
-(smaller factor).
+Assume that all subtables have $s=m/T$ cells. When $\alpha\cdot n >
+m+2s$ We can grow the first subtable while keeping the size
+constraint.  Doubling the size of a subtable increases the global
+number of cells from $m_{old} = T\cdot s$ to $m_{new} = m_{old}+s =
+(T+1)\cdot s$ (factor: $\frac{T+1}{T}$).  Note that all subsequent
+grows migrate one of the smaller tables, until all tables have the
+same size again.  Therefore, each grow until then increase the
+capacity by the same absolute amount (smaller factor).
 
@@
 
-Shrinking can work similarly to growing. We replace a second level
-table with a smaller one by migrating elements from one to the
-other.  During this migration we join elements from two buckets into
-one. Therefore, it is possible for a bucket to overfill.  We reinsert
-these elements at the end of the migration.  Obviously, this can only
-apply to at most half the migrated elements.
+Shrinking can work similarly to growing. We replace a subtable with a
+smaller one by migrating elements from one to the other.  During this
+migration we join elements from two buckets into one. Therefore, it is
+possible for a bucket to overfill.  We reinsert these elements at the
+end of the migration.  Obviously, this can only apply to at most half
+the migrated elements.
 
@@
 bits. They can address up to $2^{32}$ buckets which achieves a
-capacity for \(2^{35} \approx 34\) billion elements (bucket size 8),
+capacity for $2^{35} \approx 34$ billion elements (bucket size 8),
 at that point a hash table consumes 512GiB memory.
@@
 \label{sec:inhom_res}
-By growing individual second level tables we introduce a size imbalance
+By growing subtables individualy we introduce a size imbalance
 between subtables.  Large tables contain more buckets, but the number
@@
 
-If there are $n$ elements in a table with \emph{tl} second level tables,
+If there are $n$ elements in a hash table with $T$ subtables,
 $j$ of which have size $2s$ the others have size $s$. If elements are
 spread equally among buckets then all small tables have around
-$n/(tl+j)$ elements, and the bigger tables have $2n/(tl+j)$ elements.
-For each table there are about $kn/tl$ elements associated to one of
+$n/(T+j)$ elements, and the bigger tables have $2n/(T+j)$ elements.
+For each table there are about $Hn/T$ elements associated to one of
 its buckets.  This shows that having more hash functions can lead to a
@@
 
-For two hash functions ($k=2$) and only one grown table ($j=1$) this
-means that $\approx 2n/(tl+1)$ should be stored in the first table.
+For two hash functions ($H=2$) and only one grown table ($j=1$) this
+means that $\approx 2n/(T+1)$ should be stored in the first table.
 These are nearly all elements associated to a bucket in the first
-table ($\approx 2n/tl$). So to distribute elements evenly nearly all
+table ($\approx 2n/T$). So to distribute elements evenly nearly all
 elements would have to be stored there.
@@
 
-\subsection{Comparison Implementations}
+\subsection{Methodology and Comparison Implementations}
+
+\subsubsection*{Choice of Constants}
+In our tests, we use three hash functions ($H=3$) and a bucket size of
+($B=8$). These values have consistently outperformed other
+possibilities.  We chose to use 256 subtables for all our tests
+($T=256$).  The number of subtables only plays a minor role for the
+running time.
+
+\subsubsection*{Comparison Implementations}
 To put the performance of our solution into perspective, we
@@
 
-\textbf{Linear Probing} is the most common hashing technique outside of
+\paragraph*{Linear Probing} is the most common hashing technique outside of
 hashing with chaining.  It is commonly accepted to be inefficient for
@@
 
-\textbf{Hopscotch Hashing} is a technique similar to linear probing where
+\paragraph*{Hopscotch Hashing} is a technique similar to linear probing where
 elements are moved when they are too far from their original hash
@@
 
-\textbf{Robin Hood Hashing} is another technique based on linear probing. The
+\paragraph*{Robin Hood Hashing} is another technique based on linear probing. The
 goal of this technique is to improve find performance by minimizing
@@
 
-\textbf{Classic Bucket Cuckoo Hashing} is the technique we adopted to design
-our hashing technique.  It has clear advantages for space efficient
-tables.  Like our variant it offers constant lookup and deletion
-times.  In contrast to our variant, it does not suffer from any
-imbalances caused either by table resolution or by local growing
+\paragraph*{Classic Bucket Cuckoo Hashing} is the technique we adopted to
+design our hashing technique.  It has clear advantages for space
+efficient tables.  Like our variant it offers constant lookup and
+deletion times.  In contrast to our variant, it does not suffer from
+any imbalances caused either by table resolution or by local growing
 operations.  Growing works similar to the linear probing table above.
+Similar to DySECT, $H=3$ and $B=8$ have turned out to be an optimum
+for performance.
 
-\subsection{Hardware}
+\subsubsection*{Hardware}
 We test our implementation on two very different systems, first a
@@
 
-\subsection{Tests with a Static Table Size}
+\subsection{Influence of Fill Ratio under a Static Table Size}
+\label{sec:exp_eps}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../eval/plots/eps_insert.pdf}
+  \caption{\label{fig:eps_insert}Insert timings at different fill
+    degrees into a non-growing table.}
+\end{figure}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.49\textwidth]{../eval/plots/eps_find_s.pdf}
+  \includegraphics[width=0.49\textwidth]{../eval/plots/eps_find_u.pdf}
+  \caption{Timings for find operations with different fill degrees
+    (\emph{left:}successful, \emph{right:}unsuccessful).}
+  \label{fig:eps_find}
+\end{figure}
 Our hash table is designed for space efficient growing, therefore, its
@@
 
-DySECT fails earlier than cuckoo hashing.  The cause for this is
-twofold. First, our architecture ($T$ subtables with power of two
-sizes) does not allow exactly $24M$ cells, therefore, our table is
-slightly more full. Secondly, to get close to $24M$ cells, we have
-$110$ large subtables (twice the size of the others).  This introduces
-inhomogeneous table resolution (see \ref{sec:inhom_res}), which
-decreases our performance compared to cuckoo hashing.  This test shows
-the influence of inhomogeneous table resolution without the influence
-of size changes, and the connected table imbalance.
+DySECTs performance degrades earlier than cuckoo hashing.  The cause
+for this is twofold. First, our architecture -- $T$ subtables with power
+of two sizes -- does not allow exactly
+$24M$ cells , therefore, our table is slightly more full. Secondly, to
+get close to $24M$ cells, we have $110$ large subtables (twice the
+size of the others).  This introduces inhomogeneous table resolution
+(see \ref{sec:inhom_res}), which decreases our performance compared to
+cuckoo hashing.  This test shows the influence of inhomogeneous table
+resolution without the influence of size changes, and the connected
+table imbalance.
 
@@
 
-\subsubsection*{Constant Sized Table Under Varying Fill Degree}
-\begin{figure}
-  \includegraphics[width=\textwidth]{../eval/plots/eps_insert.pdf}
-  \caption{Insert timings at different fill degrees into a non-growing table.}
-  \label{fig:eps_insert}
+\subsection{Constructing a Table with Growing}
+\label{sec:exp_ti}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
+  \caption{\label{fig:ti_insert}Average time per insert operation
+    during the construction of a dynamic table.}
 \end{figure}
-\begin{figure}
-  \includegraphics[width=0.48\textwidth]{../eval/plots/eps_find_s.pdf}
-  \includegraphics[width=0.48\textwidth]{../eval/plots/eps_find_u.pdf}
-  \caption{Timings for find operations with different fill degrees (\emph{left:}successful, \emph{right:}unsuccessful).}
-  \label{fig:eps_find}
+%% \begin{figure}
+%%   \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_s.pdf}
+%%   \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_u.pdf}
+%%   \caption{Timings for find operations on the grown table (from Figure~\ref{fig:ti_insert}; \emph{left:}successful, \emph{right:}unsuccessful).}
+%%   \label{fig:ti_find}
+%% \end{figure}
+In this test $20M$ elements are inserted into a previously empty
+table. All tables are initialized expecting 50\,000 elements, thus
+growing is necessary, to fit all elements.  The tables are configured,
+to guarantee a load factor of
+$\varepsilon$. Figure~\ref{fig:ti_insert} shows the performance in
+relation to $\varepsilon$.  Insertion times are normalized similar to
+Figure~\ref{fig:eps_insert} (divided by $\frac{1}{\varepsilon}$) to
+make them more readable.
+
+We see that DySECT performs by far the best even with less filled
+tables.  Here we achieve a speedup of TODO$\%$ over the next best
+solution (blabla vs. linear probing blabla at $\varepsilon = 0.$ bla).
+On denser instances, we can increase this margin to TODO$\%$ (blabla
+vs we'll see blabla at $\varepsilon = 0.$ bla).  With shrinking
+$\varepsilon$, we see that insertion times start to degrade for our
+competitors.  DySECT remains close to $O(\frac{1}/{\varepsilon})$ even
+for fill degrees up to 97.5$\%$ (TODO check).
+
+We also measured the performance of find operations on the created
+tables, they are similar to the performance on the static table in
+Section~\ref{sec:exp_eps} (see Figure~\ref{fig:eps_find}), therefore,
+we omit plotting them for space reasons.
+
+\subsection{Word Count a Practical use Case}
+One of the most common use cases for hash tables -- dynamic hash
+tables in particular -- is to unify data elements with the same key.
+Oftentimes it is unclear how many unique keys are present (how large
+the table will get).  In these cases, only dynamic hash tables can
+remain space efficient.  In the following test, we take the first
+block of the common crawl data-set (TODO REFERENCE) and compute a word
+count of the contained words.  The chosen block contains around 240M
+words, with around 20M unique words.
+
+For the test, we hash each word and insert it together with a counter.
+Subsequent insertions of a reocurring word increase this counter.
+Similar to the growing benchmark, we start with an empty table
+initialized for 50\,000 elements.  The performance results can be seen
+in Figure~\ref{fig:crawl}.  We do not use any normalization since --
+on average -- each word is repeated 12 times, therefore, most inserts
+will actually behave like successful find operations (returning an
+iterator to the contained element).  Our DySECT table manages to
+dominate the field, similar to the growing test in
+Section~\ref{sec:exp_ti}.  This is a hint to the fact, that insert
+performance is important even in find intensive workloads.  This
+insight is intensified in the following test
+(Section~\ref{sec:exp_mix}).
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../eval/plots/crawl.pdf}
+  \caption{\label{fig:crawl} }
 \end{figure}
 
-\subsubsection*{Constructing a Table with Growing}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
-  \caption{Average time per insert operation during the construction of a dynamic table.}
-  \label{fig:ti_insert}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_s.pdf}
-  \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_u.pdf}
-  \caption{Timings for find operations on the grown table (from Figure~\ref{fig:ti_insert}; \emph{left:}successful, \emph{right:}unsuccessful).}
-  \label{fig:ti_find}
-\end{figure}
+\subsection{Combining insert and find operations}
+In this test, we simulate an application in progress.  We use tables
+with a minimum load factor of $\varepsilon = 0.9xx$ (TODO choose one).
+The test starts a decently filled table (TODO M elements).  On this
+table we perform a mixture of insert and find operations.  Even on
 

### 2017-04-11 source/crawl_test.cpp
Laptop - read epsilon from command line

@@ -150,6 +150,8 @@ int main(int argn, char** argc)
     const size_t      cap   = c.intArg("-cap"  , 1000);
     const size_t      steps = c.intArg("-steps", 512);
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     const std::string inf   = c.strArg("-in"   , "/home/maier/WorkEnv/tobias.org");
     const std::string outf  = c.strArg("-out"  , "");

### 2017-04-11 source/del_test.cpp

@@ -156,6 +156,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps"  , -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, win, cap, steps, alpha, name);
 }

### 2017-04-11 source/eps_test.cpp

@@ -186,6 +186,9 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps"   , 512);
     const std::string name  = c.strArg("-out"     , "");
     const double      alpha = c.doubleArg("-alpha", 1.2);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
+
     const double      alstp = c.doubleArg("-alsteps", 0.005);
     const size_t      win   = c.intArg("-win"     , 1000);
 

### 2017-04-11 source/in_test.cpp

@@ -139,6 +139,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "temp");
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test, hist_count>(c, n, pre, cap, steps, alpha, name);
 }

### 2017-04-11 source/mix_test.cpp

@@ -170,6 +170,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, n0, cap, pattern, steps, alpha, name);
 }

### 2017-04-11 source/reg_test.cpp

@@ -131,6 +131,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, cap, steps, alpha, name);
 }

### 2017-04-11 source/step_test.cpp

@@ -227,6 +227,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "temp");
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,hist_count> (c, n, cap, steps, alpha, name);
 }

### 2017-04-11 source/time_test.cpp

@@ -178,6 +178,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
+    const double      eps   = c.doubleArg("-eps", -1.0);
+    if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, n0, cap, steps, alpha, name);
 }

### 2017-04-12 source/crawl_test.cpp
Laptop - changed tests to load factor (espec. eps_test)

@@ -150,7 +150,8 @@ int main(int argn, char** argc)
     const size_t      cap   = c.intArg("-cap"  , 1000);
     const size_t      steps = c.intArg("-steps", 512);
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps", -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     const std::string inf   = c.strArg("-in"   , "/home/maier/WorkEnv/tobias.org");

### 2017-04-12 source/del_test.cpp

@@ -156,7 +156,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps"  , -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, win, cap, steps, alpha, name);

### 2017-04-12 source/eps_test.cpp

@@ -38,7 +38,7 @@ struct Test
         Table::print_init_header(out);
         print(out, "n_step" , 9);
         print(out, "win"    , 5);
-        print(out, "alpha"  , 6);
+        print(out, "load"   , 6);
         print(out, "t_in"   , 8);
         print(out, "t_fip"  , 8);
         print(out, "t_fin"  , 8);
@@ -51,29 +51,29 @@ struct Test
 
     inline void print_timing(std::ostream& out,
                              size_t i,
-                             size_t n,
+                             size_t ncurr,
                              size_t win,
-                             double alpha,
+                             double load,
                              double tin,
                              double tfip,
                              double tfin,
                              size_t in_err,
                              size_t fi_err,
-                             size_t keysize,
+                             size_t n,
                              Table& table)
     {
         print(out, i     , 4);
         table.print_init_data(out);
-        print(out, n     , 9);
+        print(out, ncurr , 9);
         print(out, win   , 5);
-        print(out, alpha , 6);
+        print(out, load  , 6);
         print(out, tin   , 8);
         print(out, tfip  , 8);
         print(out, tfin  , 8);
         print(out, in_err, 6);
         print(out, fi_err, 6);
         #ifdef MALLOC_COUNT
-        print(out, malloc_count_current() - keysize*8, 7);
+        print(out, malloc_count_current() - n*8 -win*16, 7);
         #endif
         out << std::endl;
     }
@@ -94,12 +94,11 @@ struct Test
     }
 
     int operator()(size_t it, size_t n, size_t steps,
-                   double alpha, double alstp, size_t win, std::string name)
+                   double eps, double epstp, size_t win, std::string name)
     {
         constexpr size_t range = (1ull<<63) -1;
 
-        size_t keysize = double(n)*alpha;
-        size_t* inkeys = new size_t[keysize];
+        size_t* inkeys = new size_t[n];
         size_t* fikeys = new size_t[win<<1];
 
         std::uniform_int_distribution<uint64_t> dis(1,range);
@@ -116,22 +115,23 @@ struct Test
         {
             // prepare execution randomize each iteration seperately
             // to take variance into account
-            for (size_t i = 0; i < keysize; ++i)
[…]

### 2017-04-12 source/in_test.cpp

@@ -139,7 +139,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "temp");
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps", -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test, hist_count>(c, n, pre, cap, steps, alpha, name);

### 2017-04-12 source/mix_test.cpp

@@ -170,7 +170,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps", -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, n0, cap, pattern, steps, alpha, name);

### 2017-04-12 source/reg_test.cpp

@@ -131,7 +131,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps", -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, cap, steps, alpha, name);

### 2017-04-12 source/step_test.cpp

@@ -227,7 +227,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "temp");
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps", -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,hist_count> (c, n, cap, steps, alpha, name);

### 2017-04-12 source/time_test.cpp

@@ -150,15 +150,16 @@ struct Test
             print_timing(*file, i, alpha, cap, n0, n, d_in0, d_in1, d_fn0, d_fn1,
                          in_errors, fin_errors, table);
 
-            size_t temp0 = 0;
-            size_t temp1 = 0;
-            for (auto it = table.begin(); it != table.end(); it++)
-            {
-                if (it->second < 100000)
-                    ++temp1;
-                ++temp0;
-            }
-            std::cout << temp1 << "/" << temp0 << std::endl;
+            // Iterator Test
+            // size_t temp0 = 0;
+            // size_t temp1 = 0;
+            // for (auto it = table.begin(); it != table.end(); it++)
+            // {
+            //     if (it->second < 100000)
+            //         ++temp1;
+            //     ++temp0;
+            // }
+            // std::cout << temp1 << "/" << temp0 << std::endl;
         }
 
         delete[] keys;
@@ -178,7 +179,8 @@ int main(int argn, char** argc)
     const size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
     const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      eps   = c.doubleArg("-eps", -1.0);
+    const double      load  = c.doubleArg("-load" , 2.0);
+    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, n0, cap, steps, alpha, name);

### 2017-04-13 source/crawl_test.cpp
Work test changes

@@ -146,12 +146,12 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      it    = c.intArg("-it"   , 5);
-    const size_t      cap   = c.intArg("-cap"  , 1000);
-    const size_t      steps = c.intArg("-steps", 512);
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    size_t      it    = c.intArg("-it"   , 5);
+    size_t      cap   = c.intArg("-cap"  , 1000);
+    size_t      steps = c.intArg("-steps", 512);
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     const std::string inf   = c.strArg("-in"   , "/home/maier/WorkEnv/tobias.org");

### 2017-04-13 source/del_test.cpp

@@ -149,15 +149,15 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      it    = c.intArg("-it"   , 5);
-    const size_t      n     = c.intArg("-n"    , 1000000);
-    const size_t      win   = c.intArg("-pre"  , n/2);
-    const size_t      cap   = c.intArg("-cap"  , win);
-    const size_t      steps = c.intArg("-steps", 512);
+    size_t      it    = c.intArg("-it"   , 5);
+    size_t      n     = c.intArg("-n"    , 1000000);
+    size_t      win   = c.intArg("-pre"  , n/2);
+    size_t      cap   = c.intArg("-cap"  , win);
+    size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, win, cap, steps, alpha, name);

### 2017-04-13 source/eps_test.cpp

@@ -181,13 +181,13 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      it    = c.intArg("-it"      , 5);
-    const size_t      n     = c.intArg("-n"       , 2000000);
-    const size_t      steps = c.intArg("-steps"   , 512);
+    size_t      it    = c.intArg("-it"      , 5);
+    size_t      n     = c.intArg("-n"       , 2000000);
+    size_t      steps = c.intArg("-steps"   , 512);
     const std::string name  = c.strArg("-out"     , "");
-    const double      alpha = c.doubleArg("-alpha", 0.1);
-    const double      load  = c.doubleArg("-load" , 1.0-(alpha-1.0)/alpha);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    double      alpha = c.doubleArg("-alpha", 0.1);
+    double      load  = c.doubleArg("-load" , 1.0-(alpha-1.0)/alpha);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
 
     const double      epstp = c.doubleArg("-epsteps", 0.005);
     const size_t      win   = c.intArg("-win"     , 1000);

### 2017-04-13 source/in_test.cpp

@@ -133,14 +133,14 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      n     = c.intArg("-n"    , 1000000);
-    const size_t      pre   = c.intArg("-pre"  , 0);
-    const size_t      cap   = c.intArg("-cap"  , 1000000);
-    const size_t      steps = c.intArg("-steps", 512);
-    const std::string name  = c.strArg("-out"  , "temp");
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    size_t      n     = c.intArg("-n"    , 1000000);
+    size_t      pre   = c.intArg("-pre"  , 0);
+    size_t      cap   = c.intArg("-cap"  , 1000000);
+    size_t      steps = c.intArg("-steps", 512);
+    std::string name  = c.strArg("-out"  , "temp");
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test, hist_count>(c, n, pre, cap, steps, alpha, name);

### 2017-04-13 source/mix_test.cpp

@@ -161,17 +161,16 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      it    = c.intArg("-it"   , 5);
-    const size_t      n     = c.intArg("-n"    , 1000000);
-    const size_t      n0    = c.intArg("-pre"  , n/4);
-    if (n0 > n) std::cout << "n0 (pre) has to be smaller than n! set n = n0 = " << n0 << std::endl;
-    const size_t      cap   = c.intArg("-cap"  , n0);
-    const size_t      pattern = c.intArg("-pattern", 5);
-    const size_t      steps = c.intArg("-steps", 512);
+    size_t      it    = c.intArg("-it"   , 5);
+    size_t      n     = c.intArg("-n"    , 1000000);
+    size_t      n0    = c.intArg("-pre"  , n/4);
+    size_t      cap   = c.intArg("-cap"  , n0);
+    size_t      pattern = c.intArg("-pattern", 5);
+    size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, n0, cap, pattern, steps, alpha, name);

### 2017-04-13 source/reg_test.cpp

@@ -125,14 +125,14 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      it    = c.intArg("-it"   , 5);
-    const size_t      n     = c.intArg("-n"    , 2000000);
-    const size_t      cap   = c.intArg("-cap"  , n);
-    const size_t      steps = c.intArg("-steps", 512);
+    size_t      it    = c.intArg("-it"   , 5);
+    size_t      n     = c.intArg("-n"    , 2000000);
+    size_t      cap   = c.intArg("-cap"  , n);
+    size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, cap, steps, alpha, name);

### 2017-04-13 source/selection.h

@@ -298,7 +298,7 @@ struct Chooser
              , Types&& ... param)
     {
         #ifdef SPECIAL_HOPSCOTCH
-        double ratio = c.doubleArg("-alpha", HopscotchConfig<>::GrowRatio_d);
+        double ratio = c.doubleArg("-rat", HopscotchConfig<>::GrowRatio_d);
         if (ratio < 1.101)
         {
             return executeNR<Functor, NS, std::ratio<11,10> >

### 2017-04-13 source/step_test.cpp

@@ -222,13 +222,13 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      n     = c.intArg("-n"    , 1000000);
-    const size_t      cap   = c.intArg("-cap"  , 10000);
-    const size_t      steps = c.intArg("-steps", 512);
+    size_t      n     = c.intArg("-n"    , 1000000);
+    size_t      cap   = c.intArg("-cap"  , 10000);
+    size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "temp");
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,hist_count> (c, n, cap, steps, alpha, name);

### 2017-04-13 source/time_test.cpp

@@ -171,16 +171,16 @@ struct Test
 int main(int argn, char** argc)
 {
     CommandLine c(argn, argc);
-    const size_t      it    = c.intArg("-it"   , 5);
-    const size_t      n     = c.intArg("-n"    , 2000000);
-    const size_t      n0    = c.intArg("-pre"  , n/2);
+    size_t      it    = c.intArg("-it"   , 5);
+    size_t      n     = c.intArg("-n"    , 2000000);
+    size_t      n0    = c.intArg("-pre"  , n/2);
     if (n0 > n) std::cout << "n0 (pre) has to be smaller than n! set n = n0 = " << n0 << std::endl;
-    const size_t      cap   = c.intArg("-cap"  , n);
-    const size_t      steps = c.intArg("-steps", 512);
+    size_t      cap   = c.intArg("-cap"  , n);
+    size_t      steps = c.intArg("-steps", 512);
     const std::string name  = c.strArg("-out"  , "");
-    const double      alpha = c.doubleArg("-alpha", 1.1);
-    const double      load  = c.doubleArg("-load" , 2.0);
-    const double      eps   = c.doubleArg("-eps"  , 1.0-load);
+    double      alpha = c.doubleArg("-alpha", 1.1);
+    double      load  = c.doubleArg("-load" , 2.0);
+    double      eps   = c.doubleArg("-eps"  , 1.0-load);
     if (eps > 0.) alpha = 1./(1.-eps);
 
     return Chooser::execute<Test,no_hist_count> (c, it, n, n0, cap, steps, alpha, name);

### 2017-04-16 paper/paper.tex
Laptop - Paper Changes

@@
 % Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Dynamic Space Efficient Cuckoo Table when doubling is too much -- a novel approach to growing!}
+\title{Dynamic Space Efficient Cuckoo Table -- DySECT when doubling is too much!} % -- a novel approach to growing!}
 \titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column
@@
 \begin{abstract}
-A lot of previous work concentrates on optimizing the speed and space
-efficiency of hash tables.  Additionally many practical
-implementations exist.  Modern open addressing techniques can fill a
-given table to more than 95\% before the performance deteriorates
-significantly.  But once the threshold is met, copying and migrating
-the whole table still seems to be the most efficient answer.
+Research on hash table data-structures has concentrated on optimizing
+the speed and space efficiency for a long time.  Modern open
+addressing techniques can fill a given table to more than 95\% before
+the performance deteriorates significantly.  But once the threshold is
+met, copying and migrating the whole table still seems to be the most
+efficient answer.
 
@@
 to increase the capacity by a significant amount -- usually, this
-means doubling the capacity.  Growing by large factor
-$\gamma$ often leads to bad space efficiency because the newly grown table
-is only filled to less than a factor of $\gamma^{-1}$.
+means doubling the capacity.  Growing by a large factor $\gamma$ leads
+to bad space efficiency because the newly grown table is only filled
+less than a factor of $\gamma^{-1}$.
 
@@
 One aspect which a lot of research focuses on is space efficiency.
-Modern space efficient hash tables work well up to loads of 98\% and
+Modern space efficient hash tables work well up to loads of 95\% and
 more. But for this to matter, programmers have to know similarly tight
@@
 This is typically not the case, one of the most ubiquitous use cases
-for hash tables is to unify data according to some key member.  To
+for hash tables is to unify data elements by their key.  To
 guarantee good performance, we have to overestimate the necessary
@@
 buckets.  When all of them are full we move elements, to create new
-space.  We split one bucket cuckoo table into multiple subtables each
-of which can grow independent from all others.  Because the buckets
-associated with one element are spread over the different subtables,
-growing one subtable alleviates pressure from all other tables.
+space.  To solve the problem of efficient migration, we split one
+bucket cuckoo table into multiple subtables each of which can grow
+independently from all others.  Because the buckets associated with
+one element are spread over the different subtables, growing one
+subtable alleviates pressure from all others by allowing moves from a
+dense table to the grown table.
 
@@
 small factor, while only moving a small number of elements -- making
-it possible, to amortize these small size changes. The introduced size
-and occupancy imbalance between subtables is alleviated using
-displacement techniques common to cuckoo hashing. Using these
-techniques, makes our table work and grow efficiently with fill
-degrees around 97\% (TODO READ APPROPRIATE VALUE) .  Being
-approximately \ldots{} \% faster than common hash table techniques under the
-same circumstances.
+the size changes easy to amortize. The introduced size and occupancy
+imbalance between subtables is alleviated using displacement
+techniques common to cuckoo hashing. Using these techniques, makes our
+table work and grow efficiently with fill degrees of 95\% and more.
+Being approximately TODO \ldots{} \% faster than common hash table
+techniques under the same circumstances.
 
 \section{Motivation}
+\label{sec:motication}
 Space efficiency is a major factor for the development of new hashing
@@
 statically sized tables this only makes sense, if the number of unique
-elements can be estimated tightly.
+elements can be estimated with similar tight bounds.
 
@@
 unique words. Therefore, we construct a table large enough to hold
-these elements ($0.9^{-1}\cdot c$).  It is entirely possible that the
-specific word count instance only contains $0.7\cdot c$ unique words.
-The resulting table is therefore only filled 63\%.  It is clear that
-better static sized tables cannot solve this issue.
+these elements ($0.9^{-1}\cdot c$ cells).  It is entirely possible
+that the specific word count instance only contains $0.7\cdot c$
+unique words (or less).  The resulting table is therefore only filled
+63\%.  It is clear that better static sized tables cannot solve this
+issue.
 
@@
 the problem's domain and thus also works in cases where estimating the
-number of elements can be hard.  But to achieve space efficiency in
+number of elements can be hard.  To achieve space efficiency in
 scenarios where the final size is not known the hash table has to grow
-closely with the actual number of elements.  This is not efficiently
-possible with any of the growing techniques used currently.
+closely with the actual number of elements.  This is not efficient
+with any of the current techniques used for hashing and mgration.
 
 \section{Related Work}
-\begin{itemize}
-\item Some other engineering papers
-
-\item Some Space efficient Stuff  (theoretical)
-
-\item Some Space efficient Stuff  (practical)
-\end{itemize}
+As previously stated, there has been a lot of research on hash tables,
+since they were introduced in the 1950s.
 
@@
 ($\langle key, data \rangle$) and offers the following functions:
-\verb~insert~ stores a given key-value pair, \verb~find~ given a key returns if
-said key was stored, and if so it returns the stored value, and
-\verb~erase~ which deletes a previously inserted element (if present).
+\verb~insert~ -- stores a given key-value pair; \verb~find~ -- given a key returns if
+said key was stored, and if so it returns the stored value; and
+\verb~erase~ -- deletes a previously inserted element (if present).
 
@@
 $m$ the number of cells ($m > n$) in any given hash table.  We define
-the load factor as $load = n/m$.  Tables can usually only work
+the load factor as $load = n/m$, and we use $\epsilon = 1-load$ for
+the fraction of available cells.  Tables can usually only work
 efficiently up to a certain load factor.  Above that, operations get
-slower and/or have a possibility to fail.
-When implementing a hash table one has to decide between storing
-elements directly in the hash table -- \emph{Closed Hashing} -- or storing
-pointers to elements -- \emph{Open Hashing}. This has an immediate impact
-on the necessary memory.
+slower and/or have a possibility to fail.  When implementing a hash
+table one has to decide between storing elements directly in the hash
+table -- \emph{Closed Hashing} -- or storing pointers to elements --
+\emph{Open Hashing}. This has an immediate impact on the necessary
+memory.
 
@@
 paper. For closed hash tables, the memory efficiency is directly
-dependent on the element density.
+dependent on the element density.  To reach high fill degrees with
+closed hashing tables, we have to employ \emph{open addressing}
+techniques which means, that elements are not stored in predetermined
+cells, but can be stored in one of several options (e.g. linear
+probing, or cuckoo hashing).
 
 \subsection{$\alpha$-Space Efficient Hash Tables}
-We call a hashing technique $\alpha$-space efficient when it can
-work effectively using at most $\alpha \cdot size_{\min} + O(1)$
-memory (TODO DEFINE WORKING EFFICIENTLY FURTHER i.e. constant average
-running time or just look at experiments or inserting the last 10\%
-elements is less than inserting all others). In this case we define
-working efficiently as having average insertion times close to
-$O(\frac{m}{m-n})$ (average number of probes to hit an empty cell).
+\label{sec:pre_staticspace}
+We call a hashing technique $\alpha$-space efficient when it can work
+effectively using at most $\alpha \cdot size_{\min} + O(1)$ memory. In
+this case we define working efficiently as having average insertion
+times close to $O(\frac{1}{\varepsilon})$. This is a natural bound
+since it is the expected number of fully random probes needed to hit
+an empty cell.
 
@@
 
-In many closed hashing techniques -- i.e. linear probing, cuckoo
-hashing \ldots{} -- cells are the same size as elements. Therefore, being
+In many closed hashing techniques (e.g.~linear probing, cuckoo
+hashing) cells are the same size as elements. Therefore, being
 $\alpha$ space efficient is the same as operating with a load factor
-of $\alpha^{-1} = n/m$ ($\alpha\cdot n \cdot size(element) = m\cdot
-size(element)$).  This is not necessarily the case for every kind of
-hashing technique.  Some techniques like hopscotch hashing use
-additional per cell information.  This would also be the case if we
-store some per bucket data like the number of stored elements in our
-bucket cuckoo hash table.
+of $\alpha^{-1} = n/m$ (because $\alpha\cdot n \cdot size(element) = m\cdot
+size(element)$).  Because of this, we will mostly talk about the load
+factor of a table instead of its memory usage.
 
-\subsubsection*{$\alpha$-Space Efficiency for Dynamic Tables}
-The definition of a space efficient hashing technique given in (TODO
-reference) the previous section is specifically targeted for
-statically sized hash tables.  As detailed in Motivation (TODO
-REFERENCE) our goal is to construct a hash table which can efficiently
-(and tightly) grow with the number of inserted elements. Therefore, we
-want to define the notion of an $\alpha$-space efficient dynamic
-table.
+Some techniques like hopscotch hashing use additional per cell
+information.  These hash table architectures, have to be filled more
+densely to achieve the same space efficiency.
+
+\paragraph*{$\alpha$-Space Efficiency for Dynamic Tables}
+The definition of a space efficient hashing technique given aboce is
+specifically targeted for statically sized hash tables.  As detailed
+in Section~\ref{sec:motivation} our goal is to construct a hash table
+which can efficiently (and tightly) grow with the number of inserted
+elements. Therefore, we want to define the notion of an $\alpha$-space
+efficient dynamic table.
 
@@
 
-One point of discussion while defining dynamic space efficiency is the
-space consumption during hash table operations.  During a table
-migration both the source ($t_{source}$) and the target
-($t_{target}$) table are allocated, thus using $mem =
-(m_{source}+m_{target})\cdot size(cell)$ space.  This is especially
-prohibiting if all elements are held in one table, then at least
-$2\cdot m\cdot size(element)$ is necessary for each migration.
+One problem for many implementations of space efficient hash tables is
+the migration.  During a normal full table migration, both the
+original table, and the new table are allocated.  This takes
+$\textit{m}_{\textit{new}} + \textit{m}_{\textit{old}}$
+cells. Therefore, no normal full table migration can be more then
+2-space efficient.  The only option to perform a full table migration
+with less memory, is to increase the memory in place (see
+Section~\ref{sec:dys_inplace} and
+Section~\ref{sec:exp_competitor}).
 
-Our definition of dynamic space efficiency does not enforce size
-reduction due to the deletion of elements, but it enforces that the
-memory used by deleted elements must be reused (no tombstones).
-
-\paragraph*{Trivial Examples}
-One simple example of this is a linear probing table, which grows by a
-factor of two, whenever 50\% of the table is filled. Here the table is
-at least 25\% filled (immediately after growing), therefore, the table
-is 6-space efficient (during the grow operation there are $(4+2)+n$
-cells). Another example would be a cuckoo table which doubles in size
-whenever it surpasses 95\%, fill rate. This would be a
-$0.95^{-1}+0.45^{-1}\approx 3.2$-space efficient table.
+Similar to static $\alpha$-space efficiency, we will mostly talk about
+the \emph{minimum load factor}, instead of $\alpha$.  It should be
+noted, that this definition does not enforce size reduction due to the
+deletion of elements, but it enforces that the cells used by deleted
+elements must be reused.
 
 \subsection{Cuckoo Hashing}
-Cuckoo hashing is a technique to resolve hash conflicts in an closed
+Cuckoo hashing is a technique to resolve hash conflicts in a closed
 hashing hash table. Its main draw is that it guarantees constant
@@
 positions are occupied one can often move elements to create space for
-the current element. We call this displacing elements.
+the current element. We call this process displacing elements.
 
@@
 appointed to one bucket can be stored in any of the bucket's cells .
-Using buckets one can increase the number of displacement
-opportunities drastically.
+Using buckets one can drastically increase the number of displacement
+opportunities.
 
-\textbf{Lookup:} Find and Remove operations have a guaranteed constant
+\emph{Find} and \emph{remove} operations have a guaranteed constant
 running time. Independent from the tables density, there are $H$
 possible buckets -- $H\cdot B$ cells -- that have to be searched to
-find an element.
+find an element (elements can be removed easily).
 
-\textbf{Insert:} Each element is hashed to $H$ buckets.  If at least
+During an \emph{insert} the element is hashed to $H$ buckets.  If at least
 one of those buckets has space left, then we store the element in the
@@
 
+\label{sec:cuckoo_graph}
 To visualize the problem of displacing elements, one can think of the
@@
 
-\subsubsection*{Some Performance Bounds}
 \begin{itemize}
+\item copy some performance bounds from a cuckoo hashing paper
 \item bla how dense can we get
@@
 
-When randomly probing the table we expect to find a free cell after $\frac{1}{1-load factor}$
+When randomly probing the table we expect to find a free cell after
+$\frac{1}{\varepsilon}$ random probes.
 
@@
 remains space efficient at all times.  A commonly used growing
-technique is to increase the size by migrating all elements
+technique is to double the size by migrating all elements
 into a larger new table.  Of course this is
@@
 in different subtables.  Each inserted element lies in one of its
-associated buckets.  To find a bucket associated with an
-element $e$, we compute $e$s hash value using the appropriated hash function
-$h_i(e)$. We split that hash value into two parts, then we use the
-first part to chose a subtable, and the second part to
-choose one of the contained buckets.
+associated buckets.  To find a bucket associated with an element $e$,
+we compute $e$'s hash value using the appropriate hash function
+$h_i(e)$, from a hashed key, we have to compute the subtable and the bucket
+within that subtable.
+
+To make this efficient we use powers of two for the number of
+subtables ($T = 2^t$), as well as the individual subtable sizes ($s =
+2^x$).  Since the number of subtables remains constant we can use the
+first $t$ bits from the hashed key, to find the appropriate subtable.
+From the remaining bits we compute the offset into this subtable using
+a bitmask ($h_i(e)~\texttt{AND}~s-1 \leftrightarrow h_i(e) \mod s$).
+
+%% We split that hash value into two parts, then we use the
+%% first part to chose a subtable, and the second part to
+%% choose one of the contained buckets.
 
@@
 \subsection{Growing}
-When the table contains enough elements, that the memory constraint
-can be kept, we migrate a subtable.  We migrate subtables in order
+We allocate grow one subtable,
+when the table contains enough elements, that the memory constraint
+can be kept during the migration.  We migrate subtables in order
 from the first to the last, therefore, no subtable can be more than
@@
 Assume that all subtables have $s=m/T$ cells. When $\alpha\cdot n >
-m+2s$ We can grow the first subtable while keeping the size
+m+2s$ we can grow the first subtable while keeping the size
 constraint.  Doubling the size of a subtable increases the global
@@
 
-From a theoretical point of view, the cost of growing the subtable is
-amortized by element insertions.  There are at least $\alpha^{-1}
-\cdot s = \Omega(s)$ insertions between two migrations.  One migration
-takes $\Theta(s)$ time. Alternatively abstract point of view, after
-inserting $\alpha^{-1} \cdot s$ elements the table grows enough, to
-raise the capacity about the same $\alpha^{-1} \cdot s$ effective
-cells (actual cells $\times$ fill factor).
+The cost of growing the subtable is amortized by all insertions that
+happened since the last subtable migration.  There are at least
+$\alpha^{-1} \cdot s = \Omega(s)$ insertions between two migrations.
+One migration takes $\Theta(s)$ time.  Appart from being amortized,
+the migration is efficient in practise because it accesses cells in a
+linear fashion making it really cache efficient.  Even in the target
+table cells are accessed linearly.  Because of the way we assign
+elements to buckets, one original bucket is split into two target
+buckets.  Therefore, no displacements are necessary, since no bucket
+can be overfilled.
 
-From a practical view, the migration is efficient because it accesses
-cells in a linear fashion making it really cache efficient.  Even in
-the target table cells are accessed linearly, because of the way we
-assign elements to buckets there are no displacements necessary.  The
-elements from each original bucket are split into two buckets of the
-target table.  Therefore, no bucket of the target table can have more
-elements than its respective original bucket.
-
-In the implicit graph model of the cuckoo table (TODOsee section
-cuckoo hashing), growing a subtable is equivalent to
+In the implicit graph model of the cuckoo table
+(Section~\ref{sec:cuckoo_graph}), growing a subtable is equivalent to
 splitting each node that represents a bucket within that table. The
@@
 follow a shrink operation.  When shrinking is enabled we propose to
-shrink one subtable when $\alpha*n < m-s'$ elements ($s'$ size of a
+shrink one subtable when $\alpha\cdot n < m-s'$ elements ($s'$ size of a
 large table, $m_{new} = m_{old} -s'/2$).
 
-Notice that the memory requirement actually increases during the
-operation itself.
-\subsection{Hashing Elements to Buckets}
-\subsubsection*{Compute the Bucket from a Hashed Key}
-From a hashed key, we have to compute the subtable and the bucket
-within that subtable.  To make this efficient we use powers of two for
-the number of subtables ($T = 2^t$), as well as the individual
-subtable sizes ($s = 2^*$).  Since the number of subtables remains
-constant we can use the first $t$ bits from the hashed key, to find the
-appropriate subtable.  From the remaining bits we compute the offset
-into this subtable using a bitmask ($\texttt{AND}~s-1 \leftrightarrow
-\mod s$).
+Notice that the allocated memory actually increases during the
+migration itself.
+
+\subsection{Implementation Details}
+For our experiments in Section~\ref{sec:exp} we use three hash
+functions ($H=3$) and a bucket size of ($B=8$). These values have
+consistently outperformed other possibilities.  We chose to use
+$T=256$ subtables for all our tests.  The number of subtables only
+plays a minor role for the running time.  We use breadth first search
+as displacement strategy. Breadth first search performs better in
+bucket cuckoo hashing, since it better uses the read cache lines from
+one bucket.
 
@@
 the table.  The hash function we use computes 64bit hash values
-(i.e. xxHash TODO).  We split these 64bit value into two 32bit hash
+(i.e. xxHash TODO).  We split the 64bit value into two 32bit hash
 values. All common bucket hash table sizes can be addressed using 32
@@
 buckets.
+
+\subsubsection*{Table Architecture (two Variants)}
+\label{sec:dys_inplace}
+The hash table can be implemented as a collection of pointers to
+subtables. We have to lookup the corresponding pointer, whenever a
+subtable is accessed.  This does not impact performance much since all
+subtable pointers will usually be cached.
+
+Another option, which offers some benefits is to use virtual memory,
+and memory overcommiting.  The idea is that the operating system will
+usually allow larger memory allocations than the machine has memory,
+with the anticipation that not all the allocated memory will actually
+be used.  Only memory pages, that are actually used will be mapped
+from virtual to physical memory pages.  Thus from a strict reasource
+point of view, the memory is not yet used.  Therefore, we can also
+implement the subtables as sections within one huge allocation.  This
+has the advantage, that the offset for each table can be quickly
+computed, without looking it up from a table.
+
+The added advantage is that we can grow subtables in-place. To map
+virtual to physical memory we have to initialize part of the table. To
+increase the size of a subtable, it is enough to initialize a new
+section of the table.  Once this is done, we have to redistribute all
+elements between the two parts of the table (moving only half the
+elements).  This allows us to grow a subtable without the space
+overhead of reallocation.
+
+This way, we can also implement dynamic space efficient variants of
+several other hashing schemes (see Section~\ref{sec:exp_competitors}).
+These implementations, can easily compete with similar implementations
+that use reallocation.
 
@@
 For two hash functions ($H=2$) and only one grown table ($j=1$) this
-means that $\approx 2n/(T+1)$ should be stored in the first table.
-These are nearly all elements associated to a bucket in the first
-table ($\approx 2n/T$). So to distribute elements evenly nearly all
-elements would have to be stored there.
+means that $\approx 2n/(T+1)$ should be stored in the first table, to
+achieve a balanced bucket distribution.  Therefore, nearly all
+elements associated to a bucket in the first table ($\approx 2n/T$)
+have to be stored there.
 
@@
 
-Assume the global table is filled close to 100\% when the first table begins
-to grow.  now there is the capacity for $s$ new elements, but this
-capacity is only in the first table, elements that are not hashed to
-the first table, automatically trigger displacements leading to slow
-insertions.
-
-Notice that repeated operations help to equalize this imbalance,
-because elements are more likely inserted into the less dense areas,
-and more likely to be deleted from more dense areas.
+Assume the global table is filled close to 100\% when the first table
+begins to grow.  now there is the capacity for $s$ new elements, but
+this capacity is only in the first table, elements that are not hashed
+to the first table, automatically trigger displacements leading to
+slow insertions.  Notice that repeated operations help to equalize
+this imbalance, because elements are more likely inserted into the
+less dense areas, and more likely to be deleted from more dense areas.
 
@@
 \section{Experiments}
-As described in section (TODO REFERENCE DIFFICULTIES\ldots{}) there are
-many factors that impact hash table performance.  Many of these
-factors are inherent to the combination of growing and space
-efficiency.  To show that our ideas work from a practical point of
-view we use a wide portfolio of both micro-benchmarks and practical
-experiments.
+\label{sec:exp}
+There are many factors that impact hash table performance.  Many of
+these factors are inherent to the combination of growing and space
+efficiency.  To show that our ideas work in practice we use a
+portfolio of both micro-benchmarks and practical experiments.
 
@@
 
-\subsubsection*{Choice of Constants}
-In our tests, we use three hash functions ($H=3$) and a bucket size of
-($B=8$). These values have consistently outperformed other
-possibilities.  We chose to use 256 subtables for all our tests
-($T=256$).  The number of subtables only plays a minor role for the
-running time.
-
 \subsubsection*{Comparison Implementations}
+\label{sec:exp_competitor}
 To put the performance of our solution into perspective, we
@@
 
-To grow our linear probing table, we use table migrations.  This can
-be made space efficient, by increasing the capacity in place.  This is
-possible due to over allocation (MAYBE SOME SOURCE/APPENDIX/WRITE
-SHORT PAPER) and a simple in place migration algorithm (ALSO
-APPENDIX/SHORT PAPER).  For a table that was initialized to hold $n'$
-elements with a space-efficiency $\alpha$ we trigger growing once the
-number of elements is $n = \frac{(1+\alpha)}{2}n'$. We then increase
-the capacity $m$ to $\alpha\cdot n$ to remain strictly size-efficient.
-Our implementation of the migration is highly cache efficient, but
-all elements have to be moved.  This makes repeated migrations inefficient
-especially for small $\alpha$.
+To grow our linear probing table, we use full table migrations.  This
+can be made space efficient, by increasing the capacity in place.
+This is possible due to over allocation (similar to
+Section~\ref{sec:dys_inplace}) and a simple in place migration
+algorithm (APPENDIX OR SECTION?).  For a table that was
+initialized with $m$ cells with a min load factor $1-\varepsilon$ we
+trigger growing once the table is loaded more than
+$1-\frac{\varepsilon}{2}$. We then increase the capacity $m$ to
+$(1-\varepsilon)^{-1}\cdot n$ to remain strictly size-efficient.  Our implementation
+of the migration is highly cache efficient, but all elements have to
+be moved.  This makes repeated migrations inefficient especially for
+small $\alpha$.
 
@@
 
+TODO DESCRIBE 2 CELL BUCKET VARIANT
+
 \paragraph*{Robin Hood Hashing} is another technique based on linear probing. The
 goal of this technique is to improve find performance by minimizing
-the maximum/average displacement distance.  The drawback of this technique is
-that insertions have to compute the hash of each element that is
-skipped by the insertion.
+the maximum/average displacement distance.  The drawback of this
+technique is that insertions have to compute the hash of each element
+that is skipped by the insertion.  Migrating a Robin Hood hash table
+is very fast, since the additional hash computations are not
+necessary, and the element order remains the same (sorted by hash value).
 
 \paragraph*{Classic Bucket Cuckoo Hashing} is the technique we adopted to
-design our hashing technique.  It has clear advantages for space
-efficient tables.  Like our variant it offers constant lookup and
-deletion times.  In contrast to our variant, it does not suffer from
-any imbalances caused either by table resolution or by local growing
-operations.  Growing works similar to the linear probing table above.
+design our hashing technique.  Like our variant it offers constant
+lookup and deletion times.  But it suffers from an inefficient
+growing technique similar to the one used by the linear probing table above.
 Similar to DySECT, $H=3$ and $B=8$ have turned out to be an optimum
@@
 \label{sec:exp_eps}
+
+The following test was performed, by initializing a table with
+$m\approx20M$~cells (non-growing). Then elements are inserted, until there
+is a critical amount of failing insertions ($>100$).  At different
+stages, we measure the speed of new insertions
+(Figure~\ref{fig:eps_insert}), and finds (Figure~\ref{fig:eps_find}).
+Performance is measured by performing each operation 1000 times, finds
+are measured using either randomly selected elements from within the table (successful), or by
+searching random elements from key space (unsuccessful).
+
 \begin{figure}[ht]
@@
 \end{figure}
-Our hash table is designed for space efficient growing, therefore, its
-main advantages show in dynamic workloads.  We still performed some
-tests with a static table size.  With these measurements, we can
-highlight different points without the influence of dynamically growing.
-
-The following test was performed, by initializing a table with
-$m\approx24M$~cells (non-growing). Then elements are inserted, until there
-is a critical amount of failing insertions ($>100$).  At different
-stages, we measure the speed of new insertions
-(Figure~\ref{fig:eps_insert}), and finds (Figure~\ref{fig:eps_find}).
-Performance is measured by performing each operation 1000 times, finds
-are measured using either random elements within the table (successful), or by
-searching random elements from key space (unsuccessful).
 
@@
 normalized with $\frac{1}{\varepsilon}$, which is the expected number
-of fully random probes, to find a free cell. We see that -- up to a
-certain point -- for all tables the insertion time behaves
-proportional to $\frac{1}{\varepsilon}$, then it begins to decrease
-uncontrolled.  The plot shows, that our table remains stable for
-longer than linear probing variants, but not as long, as for classic
-cuckoo hashing.  Hopscotch hashing fails earlier than all other
-tables, because its bitmaps become to small, to store the neighborhood
-information.
+of fully random probes, to find a free cell and thus a natural
+estimate for the running time. We see that -- up to a certain point --
+for all tables the insertion time behaves proportional to
+$\frac{1}{\varepsilon}$, then it begins to decrease uncontrolled.  The
+plot shows, that our table remains stable for longer than linear
+probing variants, but not as long, as for classic cuckoo hashing.
+Hopscotch hashing fails earlier than all other tables, because its
+bitmaps become to small, to store the neighborhood information.
 
-DySECTs performance degrades earlier than cuckoo hashing.  The cause
-for this is twofold. First, our architecture -- $T$ subtables with power
-of two sizes -- does not allow exactly
-$24M$ cells , therefore, our table is slightly more full. Secondly, to
-get close to $24M$ cells, we have $110$ large subtables (twice the
-size of the others).  This introduces inhomogeneous table resolution
-(see \ref{sec:inhom_res}), which decreases our performance compared to
-cuckoo hashing.  This test shows the influence of inhomogeneous table
-resolution without the influence of size changes, and the connected
-table imbalance.
+DySECTs performance degrades earlier than cuckoo hashing.  One reason
+for this is that our hash table consists of $T$ subtables with power
+of two sizes and thus does not allow exactly $20M$ cells.  Therefore,
+our table is slightly more full.  Another reason is that to get close
+to $20M$ cells, we have TODO ??? large subtables (twice the size of
+the others).  This introduces inhomogeneous table resolution (see
+\ref{sec:inhom_res}), which decreases our performance compared to
+cuckoo hashing.
 
@@
 \label{sec:exp_ti}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
-  \caption{\label{fig:ti_insert}Average time per insert operation
-    during the construction of a dynamic table.}
-\end{figure}
 %% \begin{figure}
@@
 In this test $20M$ elements are inserted into a previously empty
-table. All tables are initialized expecting 50\,000 elements, thus
+table.  The table is initialized expecting 50\,000 elements, thus
 growing is necessary, to fit all elements.  The tables are configured,
-to guarantee a load factor of
-$\varepsilon$. Figure~\ref{fig:ti_insert} shows the performance in
-relation to $\varepsilon$.  Insertion times are normalized similar to
+to guarantee a load factor of at least
+$1-\varepsilon$. Figure~\ref{fig:ti_insert} shows the performance in
+relation to the load factor.  Insertion times are normalized similar to
 Figure~\ref{fig:eps_insert} (divided by $\frac{1}{\varepsilon}$) to
 make them more readable.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
+  \caption{\label{fig:ti_insert}Average time per insert operation
+    during the construction of a dynamic table.}
+\end{figure}
 
@@
 One of the most common use cases for hash tables -- dynamic hash
-tables in particular -- is to unify data elements with the same key.
-Oftentimes it is unclear how many unique keys are present (how large
-the table will get).  In these cases, only dynamic hash tables can
-remain space efficient.  In the following test, we take the first
+tables in particular -- is to unify data elements according to their
+key.  Oftentimes it is unclear how many unique keys there are (how
+large the table will get).  In these cases, only dynamic hash tables
+can remain space efficient.  In the following test, we take the first
 block of the common crawl data-set (TODO REFERENCE) and compute a word
 count of the contained words.  The chosen block contains around 240M
-words, with around 20M unique words.
-
-For the test, we hash each word and insert it together with a counter.
-Subsequent insertions of a reocurring word increase this counter.
-Similar to the growing benchmark, we start with an empty table
-initialized for 50\,000 elements.  The performance results can be seen
-in Figure~\ref{fig:crawl}.  We do not use any normalization since --
-on average -- each word is repeated 12 times, therefore, most inserts
-will actually behave like successful find operations (returning an
-iterator to the contained element).  Our DySECT table manages to
-dominate the field, similar to the growing test in
-Section~\ref{sec:exp_ti}.  This is a hint to the fact, that insert
-performance is important even in find intensive workloads.  This
-insight is intensified in the following test
-(Section~\ref{sec:exp_mix}).
+words, with around 20M unique words. For the test, we hash each word
+and insert it's hash together with a counter.  Subsequent insertions
+of a reocurring word increase this counter.  Similar to the growing
+benchmark, we start with an empty table initialized for 50\,000
+elements.
 
@@
 
+The performance results can be seen in Figure~\ref{fig:crawl}.  We do
+not use any normalization since -- on average -- each word is repeated
+12 times, therefore, most inserts will actually behave like successful
+find operations (returning an iterator to the contained element).  Our
+DySECT table manages to dominate the field, similar to the growing
+test in Section~\ref{sec:exp_ti}.  This is a hint to the fact, that
+insert performance is important even in find intensive workloads.
+This insight is intensified in the following test where we mix insert
+and find operations.  (Section~\ref{sec:exp_mix}).
+
 \subsection{Combining insert and find operations}
-In this test, we simulate an application in progress.  We use tables
-with a minimum load factor of $\varepsilon = 0.9xx$ (TODO choose one).
-The test starts a decently filled table (TODO M elements).  On this
-table we perform a mixture of insert and find operations.  Even on
+\label{sec:exp_mix}
+In this test, we show how the hash tables behave under mixed
+workloads.  We use tables with a minimum load factor of $\varepsilon =
+0.9$ (TODO verify).  The test starts a decently filled table
+(15M elements).  On this table we perform 10M operations mixed between inserts and
+find operations.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{../eval/plots/mix.pdf}
+  \caption{\label{fig:mix} }
+\end{figure}
+
+This test clearly shows, what the last has only hinted at, the
+performance of insertions is very important even if there are
+significantly more finds than insertions.  This is even more
+accentuated by the fact that all performed finds are successful finds,
+which have usually significantly better performance than unsuccessful
+ones (especially for linear probing; see Figure~\ref{fig:eps_find}).
+
+\subsection{Conclusion}
+\subsubsection*{Future Work}
+Our Experiments clearly show, that practical running times can be
+reached while growing dynamically.  But we were unable, to prove tight
+theoretical performance bounds.
+
+Even in practice, there remains a lot of open design space.  One could
+use a similar aproach to dynamically grow different cuckoo hashing
+variants e.g. concurrent cuckoo tables similar to Li et
+al.~\cite{AlgorithmicImprovementsForFastConcurrentCuckooHashing} or
+Nguyen and Tsigas~\cite{LockFreeCuckooHashing}; or a de-amortized
+table proposed by Kirsch and
+Mitzenmacher~\cite{UsingAQueueToDeamortizeCuckooHashingInHardware},
+and refined by Arbitman et
+al.~\cite{DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults}.
 

### 2017-04-17 paper/paper.tex
Laptop - paper changes +bibliography

@@
+% -*- mode: latex; mode:writegood mode: flyspell; ispell-local-dictionary: "en_US"; coding: utf-8 -*-
+
 \documentclass[a4paper,UKenglish]{lipics-v2016}
@@
 efficient answer.
-
-To achieve any kind of efficiency, each such migration must be
-amortized by newly inserted elements.  Therefore, each migration has
-to increase the capacity by a significant amount -- usually, this
-means doubling the capacity.  Growing by a large factor $\gamma$ leads
+Growing by a constant factor $\gamma$ often leads
 to bad space efficiency because the newly grown table is only filled
@@
 independent of $n$.
+
+We achieve this goal by splitting the hash table into subtables that
+grow independent from each other.  The cuckoo load balancing
+algorithm can move elements from one table to the other.  This makes
+sure that growing one subtable increases insertion opportunities even
+for elements that are not hashed into this one table.
  \end{abstract}
@@
 \section{Motivation}
-\label{sec:motication}
+\label{sec:motivation}
 Space efficiency is a major factor for the development of new hashing
@@
 closely with the actual number of elements.  This is not efficient
-with any of the current techniques used for hashing and mgration.
+with any of the current techniques used for hashing and migration.
 
 \section{Related Work}
-As previously stated, there has been a lot of research on hash tables,
-since they were introduced in the 1950s.
+The usage of hash tables, and other hashing based algorithms has a
+long history in computer science.  The classical methods results can
+be found in all major algorithm textbooks (TODO ...).
+
+Over the last one and a half decades, the field has regained
+attention, both from theoretical and the practical point of view.  The
+initial point of innovation was the power of two
+choices~\cite{ThePowerOfTwoChoicesInRandomizedLoadBalancing}.
+This concept led to the development of cuckoo hashing by Pagh and
+Rodler~\cite{CuckooHashing}.  This technique reopened research into
+more space efficient hash tables.  It proofed theoretically
+interesting, because probabilistic bounds for the maximum fill degree
+are highly non-trivial TODO SOME BOUNDS PAPERS.
+
+Cuckoo hashing spawned many variants, that further improve its
+performance.  The most notable one is probably bucket cuckoo hashing
+($d$-ary cuckoo hashing), a variant that groups cells together in
+buckets thus improving cache usage, and increasing displacement
+opportunities TODO BUNCH OF PAPERS.  This further improves the maximum
+fill degree. Nowadays, bucket cuckoo hashing is by far the most common
+variant. There are even further variations on bucket cuckoo hashing,
+like the generalization to continuous buckets.
+
+Further adaptations to cuckoo hashing include: a de-amortization
+technique that allows a provable worst case analysis of
+insertions~\cite{UsingAQueueToDeAmortizeCuckooHashingInHardware,
+  DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults};
+and multiple concurrent implementations either powered by bucket
+locking, transactional
+memory~\cite{AlgorithmicImprovementsForFastConcurrentCuckooHashing},
+or fully lock-less~\cite{LockFreeCuckooHashing}.
+
+Some non-cuckoo space efficient hash tables continue to use linear
+probing variants.  Robin Hood hashing is a technique that was
+originally introduced in 1985 by Celis et
+al.~\cite{RobinHoodHashing}. It has regained some popularity in
+recent years -- mainly for its interesting theoretical
+properties and the possibility to reduce the inherent variance of
+linear probing.  Hopscotch~\cite{HopscotchHashing} hashing is a
+technique, that originated in the area of concurrent hash tables, but
+it can also lead to some densely filled hash tables.  For this reason
+it is often used as a technique to achieve space efficient hash
+tables.  This is only possible in a small number of circumstances, as
+the stored bitmaps used in hopscotch hashing produce a significant
+memory overhead.
+
+All these publications show that there is a clear interest in the
+development of space efficient hash tables.  Dynamic hash tables on the
+other hand seem to be considered a solved problem.  There is an
+important paper by Dietzfelbinger at
+al.~\cite{DynamicPerfectHashingUpperAndLowerBounds} that takes on the
+problem of dynamic hash tables from a theoretical perspective.  The
+paper predates cuckoo hashing, and much of the attention for space
+efficient hashing. Thus all memory bounds presented are given without
+tight constant factors.  This lack of development is where we pick up
+and offer fast solutions that support dynamic growing with tight space
+bounds.
 
@@
 ($\langle key, data \rangle$) and offers the following functions:
-\verb~insert~ -- stores a given key-value pair; \verb~find~ -- given a key returns if
-said key was stored, and if so it returns the stored value; and
-\verb~erase~ -- deletes a previously inserted element (if present).
+\verb~insert~ -- stores a given key-value pair; \verb~find~ -- given a
+key returns if said key was stored, and if so it returns the stored
+value; and \verb~erase~ -- deletes a previously inserted element (if
+present).
 
@@
 hashing hash table. Its main draw is that it guarantees constant
-lookup times even in densely filled tables. The article TODO by TODO
-(probably mitzenmacher with or without dietzfelbinger) et al. gives a
-good overview over many cuckoo hashing variants.  The distinguishing
+lookup times even in densely filled tables.  The distinguishing
 technique of cuckoo hashing is, that $H$ hash functions ($h_1, ... ,
@@
 $\alpha^{-1} \cdot s = \Omega(s)$ insertions between two migrations.
-One migration takes $\Theta(s)$ time.  Appart from being amortized,
-the migration is efficient in practise because it accesses cells in a
+One migration takes $\Theta(s)$ time.  Apart from being amortized,
+the migration is efficient in practice because it accesses cells in a
 linear fashion making it really cache efficient.  Even in the target
@@
 be used.  Only memory pages, that are actually used will be mapped
-from virtual to physical memory pages.  Thus from a strict reasource
+from virtual to physical memory pages.  Thus from a strict resource
 point of view, the memory is not yet used.  Therefore, we can also
@@
 \label{sec:inhom_res}
-By growing subtables individualy we introduce a size imbalance
+By growing subtables individually we introduce a size imbalance
 between subtables.  Large tables contain more buckets, but the number
@@
 \subsubsection*{Imbalance through Size Changes}
-In addition to the problem of inhomogenous tables, there is an
+In addition to the problem of inhomogeneous tables, there is an
 inherent balancing problem introduced by resizing subtables. It is
@@
 
-DySECTs performance degrades earlier than cuckoo hashing.  One reason
+DySECT's performance degrades earlier than cuckoo hashing.  One reason
 for this is that our hash table consists of $T$ subtables with power
@@
 guaranteed constant running times for all find operations --
-independant of the fill degree.
+independent of the fill degree.
 
@@
 and insert it's hash together with a counter.  Subsequent insertions
-of a reocurring word increase this counter.  Similar to the growing
+of a re-occurring word increase this counter.  Similar to the growing
 benchmark, we start with an empty table initialized for 50\,000
@@
 \subsection{Conclusion}
-\subsubsection*{Future Work}
-Our Experiments clearly show, that practical running times can be
-reached while growing dynamically.  But we were unable, to prove tight
-theoretical performance bounds.
+Our Experiments have shown that our DySECT hash table is by far the
+best performing hash table on all instances that make use of its
+dynamic capabilities.  It performs up to TODO $\%$ better, than the
+next best performing table.
+
+.........
+
+But there remains much to do. We were unable, to prove tight
+theoretical performance and density bounds.
 
 Even in practice, there remains a lot of open design space.  One could
-use a similar aproach to dynamically grow different cuckoo hashing
+use a similar approach to dynamically grow different cuckoo hashing
 variants e.g. concurrent cuckoo tables similar to Li et
@@
 table proposed by Kirsch and
-Mitzenmacher~\cite{UsingAQueueToDeamortizeCuckooHashingInHardware},
+Mitzenmacher~\cite{UsingAQueueToDeAmortizeCuckooHashingInHardware},
 and refined by Arbitman et
 al.~\cite{DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults}.
+
+............
 
@@
 
-\bibliography{lipics-v2016-sample-article}
+\bibliography{paper}
 

### 2017-04-17 source/mixd_test.cpp
Laptop - mixd test

@@ -83,8 +83,11 @@ struct Test
         constexpr size_t range = (1ull<<63) -1;
 
         size_t  p_rep = n/10;
+        bool*   dbool = new bool  [pre+p_rep*pattern];
         size_t* ikeys = new size_t[pre+p_rep*pattern];
-        size_t* fkeys = new size_t[p_rep*(10-pattern)];
+        size_t* dkeys = new size_t[p_rep*(10-pattern)];
+
+        std::fill(dbool, dbool+(pre+p_rep*pattern), false);
 
         std::uniform_int_distribution<uint64_t> dis(1,range);
         std::mt19937_64 re;
@@ -95,7 +98,7 @@ struct Test
         }
         size_t pcount = 0;
         size_t icount = pre-1;
-        size_t fcount = 0;
+        size_t dcount = 0;
         for (size_t i = pre; i < p_rep*10+pre; ++i)
         {
             if (pcount < pattern)
@@ -106,14 +109,18 @@ struct Test
             {
                 std::uniform_int_distribution<uint64_t> find_dis(0,icount);
                 size_t k = find_dis(re);
-                fkeys[fcount++] = ikeys[k];
+                while (dbool[k]) k = find_dis(re);
+                dbool[k] = true;
+                dkeys[dcount++] = ikeys[k];
             }
             pcount  = (pcount < 9) ? pcount+1 : 0;
         }
+        delete[] dbool;
+
 
         std::ostream* file;
         if (name == "") file = &(std::cout);
-        else file = new std::ofstream(name + ".mix",
+        else file = new std::ofstream(name + ".mixd",
                                       std::ofstream::out | std::ofstream::app);
 
         print_headline(*file);
@@ -134,7 +141,7 @@ struct Test
             auto t1 = std::chrono::high_resolution_clock::now();
             pcount = 0;
             icount = pre;
-            fcount = 0;
+            dcount = 0;
             for (size_t i = pre; i < p_rep*10+pre; ++i)
             {
                 if (pcount < pattern)
@@ -143,8 +150,7 @@ struct Test
                 }
                 else
                 {
-                    auto temp = table.find(fkeys[fcount++]);
-                    if (temp == table.end()) ++ferrors;
+                    if (! table.erase (dkeys[dcount++])) ++ferrors;
                 }
                 pcount  = (pcount < 9) ? pcount+1 : 0;
             }
@@ -159,7 +165,7 @@ struct Test
         }
 
[…]

### 2017-04-17 source/mix_test.cpp
Laptop - small fix

@@ -139,7 +139,7 @@ struct Test
             {
                 if (pcount < pattern)
                 {
-                    if (! table.insert(ikeys[icount++], icount).second) ++errors;
+                    if (! table.insert(ikeys[icount++], i).second) ++errors;
                 }
                 else
                 {

### 2017-04-17 source/mixd_test.cpp

@@ -146,7 +146,7 @@ struct Test
             {
                 if (pcount < pattern)
                 {
-                    if (! table.insert(ikeys[icount++], icount).second) ++errors;
+                    if (! table.insert(ikeys[icount++], i).second) ++errors;
                 }
                 else
                 {

### 2017-04-18 paper/paper.tex
Work - paper (pre Sebastians stuff) plotting (v1.0) fixed ripprob

@@
 % Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Dynamic Space Efficient Cuckoo Table -- DySECT when doubling is too much!} % -- a novel approach to growing!}
+\title{DySECT -- truly space efficient dynamic hashing} % -- a novel approach to growing!}
 \titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column
@@
 \author[1]{Tobias Maier}
-\author[1]{Dr. Peter Sanders}
+\author[1]{Peter Sanders}
 \affil[1]{Karlsruhe Institute of Technology, Karlsruhe, Deutschland\\
@@
 \begin{abstract}
-Research on hash table data-structures has concentrated on optimizing
-the speed and space efficiency for a long time.  Modern open
-addressing techniques can fill a given table to more than 95\% before
-the performance deteriorates significantly.  But once the threshold is
-met, copying and migrating the whole table still seems to be the most
-efficient answer.
-Growing by a constant factor $\gamma$ often leads
-to bad space efficiency because the newly grown table is only filled
-less than a factor of $\gamma^{-1}$.
+Research has concentrated on optimizing the speed and space efficiency
+of hash tables for a long time.  Modern open addressing techniques can
+fill a given table to more than 95\% before the performance
+deteriorates significantly.  But once the threshold is met, copying
+and migrating the whole table still seems to be the most efficient
+answer.  Growing by a constant factor $\gamma$ often leads to bad
+space efficiency because the newly grown table is only filled less
+than a factor of $\gamma^{-1}$.
 
@@
 growing while keeping a strict size constraint $\alpha \cdot n$
-throughout its lifetime.  Even under these conditions, we can
-guarantee constant lookup times, and amortized insertion times,
+throughout its lifetime.  Even under these conditions, we achieve
+guaranteed constant lookup times, and amortized insertion times,
 independent of $n$.
@@
 We achieve this goal by splitting the hash table into subtables that
-grow independent from each other.  The cuckoo load balancing
-algorithm can move elements from one table to the other.  This makes
-sure that growing one subtable increases insertion opportunities even
-for elements that are not hashed into this one table.
+grow independent from each other.  The cuckoo load balancing mechanism
+can move elements from one table to the other.  This ensures that
+growing one subtable increases insertion opportunities even for
+elements that are not hashed into this one table.
  \end{abstract}
@@
 interchangeable implementations which perform well under different
-conditions is important and allows programmers
-easily adapt known solutions to new circumstances.
+conditions is important and allows programmers to easily adapt known
+solutions to new circumstances.
 
@@
 This is typically not the case, one of the most ubiquitous use cases
-for hash tables is to unify data elements by their key.  To
-guarantee good performance, we have to overestimate the necessary
-capacity whenever the exact number of unique keys is not known a
-priori.  Dynamic data-structures are necessary to alleviate this
-problem.  In circumstances where space efficiency is important, space
-efficiently growing hash tables are necessary.
+for hash tables is to unify data elements by their key.  To guarantee
+good performance, we have to overestimate the initial capacity
+whenever the exact number of unique keys is not known a priori.
+Dynamic data-structures are necessary to alleviate this problem.  In
+circumstances where space efficiency is important, we need space
+efficiently growing hash tables.
 
@@
 buckets.  When all of them are full we move elements, to create new
-space.  To solve the problem of efficient migration, we split one
-bucket cuckoo table into multiple subtables each of which can grow
-independently from all others.  Because the buckets associated with
-one element are spread over the different subtables, growing one
-subtable alleviates pressure from all others by allowing moves from a
-dense table to the grown table.
+space.  To solve the problem of efficient migration, we split the
+table into multiple subtables each of which can grow independently
+from all others.  Because the buckets associated with one element are
+spread over the different subtables, growing one subtable alleviates
+pressure from all others by allowing moves from a dense table to the
+grown table.
 
@@
 statically sized tables this only makes sense, if the number of unique
-elements can be estimated with similar tight bounds.
+elements can be estimated with similarly tight bounds.
 
@@
 unique words (or less).  The resulting table is therefore only filled
-63\%.  It is clear that better static sized tables cannot solve this
-issue.
+63\%.  It is clear that better static sized tables cannot achieve fill
+ratios greater than 70\%.
 
 Depending on the underlying problem it is impossible to tightly
-estimate the final number of elements.  Dynamic growing can be a
-solution even in these instances.  It is completely independent from
-the problem's domain and thus also works in cases where estimating the
-number of elements can be hard.  To achieve space efficiency in
+estimate the final number of elements. Since dynamic growing is independent from the problem's domain, it can be a
+solution even in these instances.  To achieve space efficiency in
 scenarios where the final size is not known the hash table has to grow
@@
 The usage of hash tables, and other hashing based algorithms has a
-long history in computer science.  The classical methods results can
-be found in all major algorithm textbooks (TODO ...).
+long history in computer science.  The classical methods and results
+can be found in all major algorithm textbooks (TODO ...).
 
@@
 attention, both from theoretical and the practical point of view.  The
-initial point of innovation was the power of two
-choices~\cite{ThePowerOfTwoChoicesInRandomizedLoadBalancing}.
-This concept led to the development of cuckoo hashing by Pagh and
-Rodler~\cite{CuckooHashing}.  This technique reopened research into
-more space efficient hash tables.  It proofed theoretically
-interesting, because probabilistic bounds for the maximum fill degree
-are highly non-trivial TODO SOME BOUNDS PAPERS.
+initial innovation that sparked this attention was the power of two
+choices~\cite{ThePowerOfTwoChoicesInRandomizedLoadBalancing}.  This
+concept led to the development of cuckoo hashing~\cite{CuckooHashing}
+which reopened research into more space efficient hash tables.  They
+proofed theoretically interesting, because probabilistic bounds for
+the maximum fill degree are often highly non-trivial TODO SOME BOUNDS
+PAPERS.
 
-Cuckoo hashing spawned many variants, that further improve its
+Cuckoo hashing also spawned many variants, that further improve its
 performance.  The most notable one is probably bucket cuckoo hashing
@@
 variant. There are even further variations on bucket cuckoo hashing,
-like the generalization to continuous buckets.
+like the generalization to continuous buckets TODO PAPER.
 
@@
 probing variants.  Robin Hood hashing is a technique that was
-originally introduced in 1985 by Celis et
-al.~\cite{RobinHoodHashing}. It has regained some popularity in
+originally introduced in 1985~\cite{RobinHoodHashing}. It has regained some popularity in
 recent years -- mainly for its interesting theoretical
@@
 linear probing.  Hopscotch~\cite{HopscotchHashing} hashing is a
-technique, that originated in the area of concurrent hash tables, but
+technique, that originated in the research of concurrent hash tables, but
 it can also lead to some densely filled hash tables.  For this reason
@@
 tables.  This is only possible in a small number of circumstances, as
-the stored bitmaps used in hopscotch hashing produce a significant
+the stored bitmaps used in hopscotch hashing create a significant
 memory overhead.
@@
 \verb~insert~ -- stores a given key-value pair; \verb~find~ -- given a
-key returns if said key was stored, and if so it returns the stored
-value; and \verb~erase~ -- deletes a previously inserted element (if
-present).
+key returns if said key was stored, and $\bot$ otherwise; and
+\verb~erase~ -- deletes a previously inserted element (if present).
 
@@
 $m$ the number of cells ($m > n$) in any given hash table.  We define
-the load factor as $load = n/m$, and we use $\epsilon = 1-load$ for
+the load factor as $load = n/m$, and we use $\varepsilon = 1-load$ for
 the fraction of available cells.  Tables can usually only work
@@
 \begin{align*}
-  mem_{closed}(n,m) &\geq^* m\cdot size(element)\\
-  mem_{open}  (n,m) &\geq^* m\cdot size(pointer) + n\cdot size(element)\\
+  mem_{closed}(n,m) &\geq m\cdot size(element)\\
+  mem_{open}  (n,m) &\geq m\cdot size(pointer) + n\cdot size(element)\\
 \end{align*}
 
-For large elements ($\gg size(pointer)$), one can simply use open
-hashing, to reduce the relevant memory factor.  Therefore, we restrict
-ourselves to the common case of elements whose size is comparable to
-that of a pointer.  For our experiments we use 128bit elements (64bit
-keys and 64bit values).  In these instances closed addressing
-introduces a significant memory overhead (at least $\times1.5$),
-therefore, we only consider closed hash tables for the purpose of this
-paper. For closed hash tables, the memory efficiency is directly
-dependent on the element density.  To reach high fill degrees with
-closed hashing tables, we have to employ \emph{open addressing}
-techniques which means, that elements are not stored in predetermined
-cells, but can be stored in one of several options (e.g. linear
-probing, or cuckoo hashing).
+For large elements ($\gg size(pointer)$), one can use a non-space
+efficient hash table with open hashing, to reduce the relevant memory
+factor.  Therefore, we restrict ourselves to the common and more interesting case of
+elements whose size is comparable to that of a pointer.  For our
+experiments we use 128bit elements (64bit keys and 64bit values).  In
+these instances open hashing introduces a significant memory
+overhead (at least $\times1.5$), therefore, we only consider closed
+hash tables for the purpose of this paper. For closed hash tables, the
+memory efficiency is directly dependent on the element density.  To
+reach high fill degrees with closed hashing tables, we have to employ
+\emph{open addressing} techniques which means, that elements are not
+stored in predetermined cells, but can be stored in one of several
+options (e.g. linear probing, or cuckoo hashing).
 
 \subsection{$\alpha$-Space Efficient Hash Tables}
+\subparagraph*{Static}
 \label{sec:pre_staticspace}
 We call a hashing technique $\alpha$-space efficient when it can work
-effectively using at most $\alpha \cdot size_{\min} + O(1)$ memory. In
+effectively using at most $\alpha \cdot n_{\textit{curr}} \cdot \textit{size}(\textit{element}) + O(1)$ memory. In
 this case we define working efficiently as having average insertion
@@
 
-\[\alpha \cdot size_min = \alpha\cdot n\cdot size(element)\]
+%\[\alpha \cdot size_min = \alpha\cdot n\cdot size(element)\]
 
@@
 $\alpha$ space efficient is the same as operating with a load factor
-of $\alpha^{-1} = n/m$ (because $\alpha\cdot n \cdot size(element) = m\cdot
-size(element)$).  Because of this, we will mostly talk about the load
-factor of a table instead of its memory usage.
+of $\alpha^{-1} = n/m$.
+%(because $\alpha\cdot n \cdot size(element) = m\cdot size(element)$).
+Because of this, we will mostly talk about the load factor of a table
+instead of its memory usage.
 
@@
 
-\paragraph*{$\alpha$-Space Efficiency for Dynamic Tables}
+\paragraph*{Dynamic}
 The definition of a space efficient hashing technique given aboce is
-specifically targeted for statically sized hash tables.  As detailed
-in Section~\ref{sec:motivation} our goal is to construct a hash table
-which can efficiently (and tightly) grow with the number of inserted
-elements. Therefore, we want to define the notion of an $\alpha$-space
-efficient dynamic table.
-
-We call a hash table implementation dynamically $\alpha$-space
-efficient if an instantiated table can grow arbitrarily large over its
-original capacity, while remaining smaller than $\alpha\cdot
-mem_{necessary}$ at all times.
-\[mem_{curr} \leq \alpha\cdot max(n_{curr})\cdot size(element)\]
+specifically targeted for statically sized hash tables.  We call a
+hash table implementation dynamically $\alpha$-space efficient if an
+instantiated table can grow arbitrarily large over its original
+capacity, while remaining smaller than $\alpha\cdot n_{\max}\cdot
+\textit{size}(\textit{element}) + O(1)$ at all times.
+%\[mem_{curr} \leq \alpha\cdot max(n_{curr})\cdot size(element)\]
 
@@
 Similar to static $\alpha$-space efficiency, we will mostly talk about
-the \emph{minimum load factor}, instead of $\alpha$.  It should be
-noted, that this definition does not enforce size reduction due to the
-deletion of elements, but it enforces that the cells used by deleted
-elements must be reused.
+the \emph{minimum load factor}, instead of $\alpha$.  We should note,
+that this definition of dynamic $\alpha$-space efficient hash tables
+does not enforce size reduction due to the deletion of elements, but
+it enforces that cells previously used by deleted elements must be reused.
 
 \subsection{Cuckoo Hashing}
+\label{sec:pre_cuckoo}
 Cuckoo hashing is a technique to resolve hash conflicts in a closed
@@
 technique of cuckoo hashing is, that $H$ hash functions ($h_1, ... ,
-h_H$) are used to find $H$ independent possible positions. Each
-element is stored in one of its positions.  Even if all possible
-positions are occupied one can often move elements to create space for
-the current element. We call this process displacing elements.
+h_H$) are used to compute $H$ independent positions. Each element is
+stored in one of its positions.  Even if all positions are occupied
+one can often move elements to create space for the current
+element. We call this process displacing elements.
 
@@
 running time. Independent from the tables density, there are $H$
-possible buckets -- $H\cdot B$ cells -- that have to be searched to
-find an element (elements can be removed easily).
+buckets -- $H\cdot B$ cells -- that have to be searched to find an
+element.
 
 During an \emph{insert} the element is hashed to $H$ buckets.  If at least
-one of those buckets has space left, then we store the element in the
-bucket, that has the most space.  But when all buckets are full, then we
+one of those buckets has an empty cell, we store the element in the
+bucket with the most space.  When all buckets are full, then we
 have to move elements around the table, such that a new space becomes
@@
 
-\begin{itemize}
-\item copy some performance bounds from a cuckoo hashing paper
-\item bla how dense can we get
-\end{itemize}
-
-When randomly probing the table we expect to find a free cell after
-$\frac{1}{\varepsilon}$ random probes.
-
 \section{DySECT Space Efficient Growing, by Splitting the Table}
@@
 into a larger new table.  Of course this is
-not memory efficient.  A natural thought which occurs is to double
+not memory efficient.  A natural thought is to double
 only part of the data structure.  This idea together with the
-balancing ability of multiple hash functions is what will achieve the
+balancing mechanism of cuckoo hashing will achieve the
 described functionality.
@@
 \subsection{Description}
-We use multiple ($T$) subtables (shown in TODO FIGREFRENCE) to allow
+We use $T$ subtables (shown in TODO FIGREFRENCE) to allow
 differentiated growing between table parts.  The subtables consist of
@@
 
-Comparable to classic cuckoo hashing, each element has $H$ associated
-buckets ($H$ number of hash functions), these can be in the same, or
-in different subtables.  Each inserted element lies in one of its
-associated buckets.  To find a bucket associated with an element $e$,
+Each element has $H$ associated
+buckets similar to cuckoo hashing, these can be in the same, or
+in different subtables.  To find a bucket associated with an element $e$,
 we compute $e$'s hash value using the appropriate hash function
@@
 subtables ($T = 2^t$), as well as the individual subtable sizes ($s =
-2^x$).  Since the number of subtables remains constant we can use the
+2^x*B$).  Since the number of subtables is constant we can use the
 first $t$ bits from the hashed key, to find the appropriate subtable.
 From the remaining bits we compute the offset into this subtable using
-a bitmask ($h_i(e)~\texttt{AND}~s-1 \leftrightarrow h_i(e) \mod s$).
+a bitmask ($h_i(e)\,\texttt{\&}\,(s-1) = h_i(e) \mod s$).
 
@@
 \subsection{Growing}
-We allocate grow one subtable,
+We grow one subtable,
 when the table contains enough elements, that the memory constraint
@@
 number of cells from $m_{old} = T\cdot s$ to $m_{new} = m_{old}+s =
-(T+1)\cdot s$ (factor: $\frac{T+1}{T}$).  Note that all subsequent
+(T+1)\cdot s$ (grow factor $\frac{T+1}{T}$).  Note that all subsequent
 grows migrate one of the smaller tables, until all tables have the
 same size again.  Therefore, each grow until then increase the
-capacity by the same absolute amount (smaller factor).
+capacity by the same absolute amount (smaller grow factor).
 
@@
 functions ($H=3$) and a bucket size of ($B=8$). These values have
-consistently outperformed other possibilities.  We chose to use
-$T=256$ subtables for all our tests.  The number of subtables only
-plays a minor role for the running time.  We use breadth first search
-as displacement strategy. Breadth first search performs better in
-bucket cuckoo hashing, since it better uses the read cache lines from
-one bucket.
+consistently outperformed other options.  We chose to use $T=256$
+subtables for all our tests.  The number of subtables only plays a
+minor role for the running time.  To find displacement opportunities
+we use breadth first search.  In our tests it performed better than
+random walks, since it better uses the read cache lines from one
+bucket.
 
@@
 
-The added advantage is that we can grow subtables in-place. To map
-virtual to physical memory we have to initialize part of the table. To
-increase the size of a subtable, it is enough to initialize a new
-section of the table.  Once this is done, we have to redistribute all
-elements between the two parts of the table (moving only half the
-elements).  This allows us to grow a subtable without the space
-overhead of reallocation.
+The added advantage is that we can grow subtables in-place.  To
+increase the size of a subtable, it is enough to initialize a
+consecutive section of the table (following the original subtable).
+Once this is done, we have to redistribute the table's elements.  This
+allows us to grow a subtable without the space overhead of
+reallocation.
 
-This way, we can also implement dynamic space efficient variants of
-several other hashing schemes (see Section~\ref{sec:exp_competitors}).
-These implementations, can easily compete with similar implementations
-that use reallocation.
+This way, we can also implement in place growing variants of other
+hashing schemes (see Section~\ref{sec:exp_competitors}).  These
+implementations have the same performance of similar implementations
+using full table migration.  But given the correct growing factor,
+they can be space efficient.
 
 \subsection{Difficulties for the Analysis of our table structure}
-There are three factors, that impact the performance of our dynamic
-table compared to other cuckoo table variants and to other hashing solutions in
-general \textbf{inhomogeneous table resolution}, \textbf{element imbalance}, and
-\textbf{population density}. All of these factors influence the maximum load
-density and the running times in different ways.
+There are two factors, that impact the performance of our dynamic
+table compared to other cuckoo table variants and to other hashing
+solutions in general \textbf{inhomogeneous table resolution} and
+\textbf{element imbalance}.
+%, and \textbf{population density}.
+All of these factors influence the maximum load density and the
+running times in different ways.
 
-\subsubsection*{Imbalance through Inhomogeneous Table Resolution}
+\subparagraph*{Imbalance through Inhomogeneous Table Resolution}
 \label{sec:inhom_res}
@@
 elements associated to a bucket in the first table ($\approx 2n/T$)
-have to be stored there.
+have to be stored there.  This is the reason, why $H=3$ leads to
+significantly better space efficiency than $H=2$.
 
-\subsubsection*{Imbalance through Size Changes}
+\subparagraph*{Imbalance through Size Changes}
 In addition to the problem of inhomogeneous tables, there is an
@@
 
-\subsubsection*{Population Density}
-Measuring and comparing the performance of space efficient growing
-tables to their statically sized counterparts is difficult.  Inserting
-elements into a densely filled table takes longer, than into an empty
-one.  Therefore, a table that is always densely filled will naturally
-be slower than one that only fills up towards the end of the
-execution -- even when ignoring eventual growing costs.
+% \subsubsection*{Population Density}
+% Measuring and comparing the performance of space efficient growing
+% tables to their statically sized counterparts is difficult.  Inserting
+% elements into a densely filled table takes longer, than into an empty
+% one.  Therefore, a table that is always densely filled will naturally
+% be slower than one that only fills up towards the end of the
+% execution -- even when ignoring eventual growing costs.
 
@@
 \subsubsection*{Comparison Implementations}
-\label{sec:exp_competitor}
+\label{sec:exp_competitors}
 To put the performance of our solution into perspective, we
-implemented and tested several other options for space efficient
-hashing.  We use our own implementations, since most hash tables found
-online do not support strict space-efficiency.  In Appendix TODO we
-show that our implementations are at least competitive with
-implementations found .
+implement and test several other options for space efficient
+hashing.  We use our own implementations, since no hash table found
+online supports strict space-efficiency.
 
-\paragraph*{Linear Probing} is the most common hashing technique outside of
+\subparagraph*{Linear Probing} is the most common hashing technique outside of
 hashing with chaining.  It is commonly accepted to be inefficient for
@@
 be moved.  This makes repeated migrations inefficient especially for
-small $\alpha$.
+small $\varepsilon$.
 
-\paragraph*{Hopscotch Hashing} is a technique similar to linear probing where
+\subparagraph*{Hopscotch Hashing} is a technique similar to linear probing where
 elements are moved when they are too far from their original hash
@@
 
-\paragraph*{Robin Hood Hashing} is another technique based on linear probing. The
+\subparagraph*{Robin Hood Hashing} is another technique based on linear probing. The
 goal of this technique is to improve find performance by minimizing
-the maximum/average displacement distance.  The drawback of this
+the maximum displacement distance.  The drawback of this
 technique is that insertions have to compute the hash of each element
@@
 
-\paragraph*{Classic Bucket Cuckoo Hashing} is the technique we adopted to
-design our hashing technique.  Like our variant it offers constant
-lookup and deletion times.  But it suffers from an inefficient
-growing technique similar to the one used by the linear probing table above.
+\subparagraph*{Classic Bucket Cuckoo Hashing} as described in
+Section~\ref{sec:pre_cuckoo} is the basic hashing technique, that
+inspired our hashing technique.  Like our variant it offers constant
+lookup and deletion times.  But it suffers from an inefficient growing
+technique similar to the one used by the linear probing table above.
 Similar to DySECT, $H=3$ and $B=8$ have turned out to be an optimum
@@
 The following test was performed, by initializing a table with
-$m\approx20M$~cells (non-growing). Then elements are inserted, until there
+$m\approx25M$~cells (non-growing). Then elements are inserted, until there
 is a critical amount of failing insertions ($>100$).  At different
@@
 for this is that our hash table consists of $T$ subtables with power
-of two sizes and thus does not allow exactly $20M$ cells.  Therefore,
+of two sizes and thus does not allow exactly $25M$ cells.  Therefore,
 our table is slightly more full.  Another reason is that to get close
-to $20M$ cells, we have TODO ??? large subtables (twice the size of
+to $25M$ cells, we have 125 large subtables (twice the size of
 the others).  This introduces inhomogeneous table resolution (see
@@
 $\varepsilon$, we see that insertion times start to degrade for our
-competitors.  DySECT remains close to $O(\frac{1}/{\varepsilon})$ even
+competitors.  DySECT remains close to $O(\frac{1}{\varepsilon})$ even
 for fill degrees up to 97.5$\%$ (TODO check).
@@
 Section~\ref{sec:exp_eps} (see Figure~\ref{fig:eps_find}), therefore,
-we omit plotting them for space reasons.
+we omit displaying them for space reasons.
 
@@
 tables in particular -- is to unify data elements according to their
-key.  Oftentimes it is unclear how many unique keys there are (how
-large the table will get).  In these cases, only dynamic hash tables
-can remain space efficient.  In the following test, we take the first
-block of the common crawl data-set (TODO REFERENCE) and compute a word
-count of the contained words.  The chosen block contains around 240M
-words, with around 20M unique words. For the test, we hash each word
-and insert it's hash together with a counter.  Subsequent insertions
-of a re-occurring word increase this counter.  Similar to the growing
-benchmark, we start with an empty table initialized for 50\,000
-elements.
+key.  In these cases, only dynamic hash tables can remain space
+efficient even though the number of unique keys is often unknown.  In
+the following test, we take the first block of the common crawl
+data-set (TODO REFERENCE) and compute a word count of the contained
+words.  The chosen block contains around 240M words, with around 20M
+unique words. For the test, we hash each word and insert it's hash
+together with a counter.  Subsequent insertions of a re-occurring word
+increase this counter.  Similar to the growing benchmark, we start
+with an empty table initialized for 50\,000 elements.
 
@@
 
-\subsection{Combining insert and find operations}
+\subsection{Access Patterns/Combining operations}
 \label{sec:exp_mix}
@@
 workloads.  We use tables with a minimum load factor of $\varepsilon =
-0.9$ (TODO verify).  The test starts a decently filled table
+0.95$ (TODO verify).  The test starts a decently filled table
 (15M elements).  On this table we perform 10M operations mixed between inserts and
-find operations.
+find/erase operations.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../eval/plots/mix.pdf}
+  \includegraphics[width=0.49\textwidth]{../eval/plots/mix.pdf}
+  \includegraphics[width=0.49\textwidth]{../eval/plots/mixd.pdf}
   \caption{\label{fig:mix} }
@@
 
-This test clearly shows, what the last has only hinted at, the
+This test clearly shows that the
 performance of insertions is very important even if there are
@@
 use a similar approach to dynamically grow different cuckoo hashing
-variants e.g. concurrent cuckoo tables similar to Li et
-al.~\cite{AlgorithmicImprovementsForFastConcurrentCuckooHashing} or
-Nguyen and Tsigas~\cite{LockFreeCuckooHashing}; or a de-amortized
-table proposed by Kirsch and
-Mitzenmacher~\cite{UsingAQueueToDeAmortizeCuckooHashingInHardware},
-and refined by Arbitman et
-al.~\cite{DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults}.
+variants e.g. different variants of concurrent cuckoo
+tables~\cite{AlgorithmicImprovementsForFastConcurrentCuckooHashing,
+  LockFreeCuckooHashing}; or a table with de-amortized
+insertions~\cite{UsingAQueueToDeAmortizeCuckooHashingInHardware,
+  DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults}.
 

### 2017-04-18 include/prob_robin.h

@@ -459,12 +459,14 @@ private:
         std::vector<value_intern> buffer;
 
         size_type distance   = 0;
-        for (int i = index; i >= 0; ++i)
+        for (int i = index; i >= 0; --i)
         {
             auto temp = table[i];
 
             if (!temp.first)
                 continue;
+
+            table[i] = value_intern();
 
             auto nind = h(temp.first);
             if (nind < size_type(i)) { buffer.push_back(temp); continue; }

### 2017-04-18 paper/paper.tex
Work - After Sebastians Improvements

@@
 \title{DySECT -- truly space efficient dynamic hashing} % -- a novel approach to growing!}
-\titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column
+\titlerunning{DySECT -- truly space efficient dynamic hashing} %optional, in case that the title is too long; the running title should fit into the top page column
 
@@
 Research has concentrated on optimizing the speed and space efficiency
-of hash tables for a long time.  Modern open addressing techniques can
-fill a given table to more than 95\% before the performance
-deteriorates significantly.  But once the threshold is met, copying
-and migrating the whole table still seems to be the most efficient
-answer.  Growing by a constant factor $\gamma$ often leads to bad
-space efficiency because the newly grown table is only filled less
-than a factor of $\gamma^{-1}$.
+of hash tables for a long time.  Hash tables using modern open
+addressing techniques can be filled to more than 95\% before the
+performance deteriorates significantly.  Once this threshold is met,
+copying and migrating the whole table still seems to be the most
+efficient answer.  Growing by a constant factor $\gamma$ often leads
+to bad space efficiency because the filling degree of the grown table
+is at most $\gamma^{-1}$. During the migration, the space efficiency
+is even worse, since oftentimes both the new and the old table coexist
+at the same time, causing a massive space overhead.
 
-The space efficiency is even worse during the migration, where
-oftentimes both the new and the old table coexist at the same time
-causing a massive space overhead.  To solve both of these issues, we
-propose the a new hash table architecture that supports dynamic
-growing while keeping a strict size constraint $\alpha \cdot n$
-throughout its lifetime.  Even under these conditions, we achieve
-guaranteed constant lookup times, and amortized insertion times,
-independent of $n$.
+We propose a new hash table
+architecture that supports dynamic growing while keeping a strict size
+constraint $\alpha \cdot n$ throughout its lifetime.  Even under these
+conditions, we achieve guaranteed constant lookup times, and amortized
+constant insertion times.
 
-We achieve this goal by splitting the hash table into subtables that
+Our architecture splits the overall hash table into subtables that
 grow independent from each other.  The cuckoo load balancing mechanism
-can move elements from one table to the other.  This ensures that
-growing one subtable increases insertion opportunities even for
-elements that are not hashed into this one table.
+can move elements between subtables ensuring that growing one
+subtable increases the overall insertion opportunities.
  \end{abstract}
@@
 \section{Introduction}
-Hash tables are some of the most frequently used data-structures. They
-belong into every programmers basic toolbox.  As such having
+Hash tables are some of the most frequently used data-structures. Having
 interchangeable implementations which perform well under different
@@
 
-One aspect which a lot of research focuses on is space efficiency.
-Modern space efficient hash tables work well up to loads of 95\% and
-more. But for this to matter, programmers have to know similarly tight
-bounds to the absolute number of elements inserted into the table.
-This is typically not the case, one of the most ubiquitous use cases
-for hash tables is to unify data elements by their key.  To guarantee
-good performance, we have to overestimate the initial capacity
-whenever the exact number of unique keys is not known a priori.
-Dynamic data-structures are necessary to alleviate this problem.  In
-circumstances where space efficiency is important, we need space
-efficiently growing hash tables.
+One aspect a lot of research focuses on is space efficiency TODO
+REFERENCES.  Modern space efficient hash tables work well up to loads
+of 95\% and more.  But for this to matter, the table has to be
+initialized with the correct final capacity, therefore, forcing
+programmers to know tight bounds to the maximum number of inserted
+elements.  This is typically not the case.  One of the most ubiquitous
+use cases for hash tables is to unify data elements by their key.
+Whenever the exact number of unique keys is not known a priori we have
+to overestimate the initial capacity to guarantee good performance.
+Dynamic space efficient data-structures are necessary to alleviate
+this problem independent of the circumstances.
+
+To visualize this assume the following scenario.  During the operation
+of a word count benchmark, we know an upper bound $c$ to the number of
+unique words. Therefore, we construct a table large enough to hold
+these elements ($0.9^{-1}\cdot c$ cells).  It is entirely possible
+that the specific word count instance only contains $0.7\cdot c$
+unique words.  The resulting table is therefore only filled to 63\%.
+It is clear that better static sized tables cannot achieve fill ratios
+greater than 70\%.  A factor of 0.7 seems realistic considering
+that the benchmark from Section~\ref{sec:exp_wordcount} for example
+contains 240\,000\,000 words only 20\,000\,000 of which are unique.
+
+Depending on the underlying problem it is impossible to tightly
+estimate the final number of elements. Since growing the table
+dynamically is independent from the problem domain, it can be a
+solution even in these instances.  To achieve space efficiency in
+scenarios where the final size is not known the hash table has to grow
+closely with the actual number of elements.  This cannot be achieved
+efficiently Swith any of the current techniques used for hashing and
+migration.
 
 Many libraries -- even ones that implement space efficient hash tables
--- already offer dynamic growing.  The problem with these
-implementations is that they either lose their space efficiency or
-their overall performance, once the table grows above its original
-capacity.  Usually growing is implemented by either creating
-additional hash tables -- decreasing the performance especially for
-look ups, or by migrating all elements to a new table -- losing the
-space efficiency by multiplying the original size.
+-- already offer growing.  The problem with these implementations is
+that they either lose their space efficiency or their performance
+degrades once the table grows above its original capacity.  Usually
+growing is implemented by either creating additional hash tables --
+decreasing the performance especially for look ups, or by migrating
+all elements to a new table -- losing the space efficiency by
+multiplying the original size.
 
@@
 each element can be stored in one of several associated constant sized
-buckets.  When all of them are full we move elements, to create new
-space.  To solve the problem of efficient migration, we split the
-table into multiple subtables each of which can grow independently
-from all others.  Because the buckets associated with one element are
-spread over the different subtables, growing one subtable alleviates
-pressure from all others by allowing moves from a dense table to the
-grown table.
+buckets.  When all of them are full we move an element into another of
+its buckets to create new space.  To solve the problem of efficient
+migration, we split the table into multiple subtables each of which
+can grow independently from all others.  Because the buckets
+associated with one element are spread over the different subtables,
+growing one subtable alleviates pressure from all others by allowing
+moves from a dense subtable to the grown subtable.
 
@@
 
-\section{Motivation}
-\label{sec:motivation}
-Space efficiency is a major factor for the development of new hashing
-techniques.  The goal has long been to make hash tables work better in
-high load scenarios.  Modern hashing techniques like Hopscotch, Robin
-Hood, and Cuckoo Hashing can easily fill hash tables to more than 90\%
-and up to 98\% while still achieving constant average running times on
-most of their operations.  But since these techniques depend on
-statically sized tables this only makes sense, if the number of unique
-elements can be estimated with similarly tight bounds.
+%% \section{Motivation}
+%% \label{sec:motivation}
+%% Space efficiency is a major factor for the development of new hashing
+%% techniques.  The goal has long been to make hash tables work better in
+%% high load scenarios.  Modern hashing techniques like Hopscotch, Robin
+%% Hood, and Cuckoo Hashing can easily fill hash tables to more than 90\%
+%% and up to 98\% while still achieving constant average running times on
+%% most of their operations.  But since these techniques depend on
+%% statically sized tables this only makes sense, if the number of unique
+%% elements can be estimated with similarly tight bounds.
 
-To visualize this assume the following scenario.  During the operation
-of a word count benchmark, we know an upper bound $c$ to the number of
-unique words. Therefore, we construct a table large enough to hold
-these elements ($0.9^{-1}\cdot c$ cells).  It is entirely possible
-that the specific word count instance only contains $0.7\cdot c$
-unique words (or less).  The resulting table is therefore only filled
-63\%.  It is clear that better static sized tables cannot achieve fill
-ratios greater than 70\%.
+%% To visualize this assume the following scenario.  During the operation
+%% of a word count benchmark, we know an upper bound $c$ to the number of
+%% unique words. Therefore, we construct a table large enough to hold
+%% these elements ($0.9^{-1}\cdot c$ cells).  It is entirely possible
+%% that the specific word count instance only contains $0.7\cdot c$
+%% unique words (or less).  The resulting table is therefore only filled
+%% 63\%.  It is clear that better static sized tables cannot achieve fill
+%% ratios greater than 70\%.
 
-Depending on the underlying problem it is impossible to tightly
-estimate the final number of elements. Since dynamic growing is independent from the problem's domain, it can be a
-solution even in these instances.  To achieve space efficiency in
-scenarios where the final size is not known the hash table has to grow
-closely with the actual number of elements.  This is not efficient
-with any of the current techniques used for hashing and migration.
+%% Depending on the underlying problem it is impossible to tightly
+%% estimate the final number of elements. Since dynamic growing is
+%% independent from the problem's domain, it can be a solution even in
+%% these instances.  To achieve space efficiency in scenarios where the
+%% final size is not known the hash table has to grow closely with the
+%% actual number of elements.  This is not efficient with any of the
+%% current techniques used for hashing and migration.
 
@@
 attention, both from theoretical and the practical point of view.  The
-initial innovation that sparked this attention was the power of two
-choices~\cite{ThePowerOfTwoChoicesInRandomizedLoadBalancing}.  This
-concept led to the development of cuckoo hashing~\cite{CuckooHashing}
-which reopened research into more space efficient hash tables.  They
-proofed theoretically interesting, because probabilistic bounds for
-the maximum fill degree are often highly non-trivial TODO SOME BOUNDS
-PAPERS.
+initial innovation that sparked this attention was the idea, that
+storing an element in the less filled of two ``random'' chains leads to
+incredibly well balanced loads.  This concept is called the power of
+two choices~\cite{ThePowerOfTwoChoicesInRandomizedLoadBalancing}.
 
-Cuckoo hashing also spawned many variants, that further improve its
-performance.  The most notable one is probably bucket cuckoo hashing
-($d$-ary cuckoo hashing), a variant that groups cells together in
-buckets thus improving cache usage, and increasing displacement
-opportunities TODO BUNCH OF PAPERS.  This further improves the maximum
-fill degree. Nowadays, bucket cuckoo hashing is by far the most common
-variant. There are even further variations on bucket cuckoo hashing,
-like the generalization to continuous buckets TODO PAPER.
+It led to the development of cuckoo hashing~\cite{CuckooHashing}.
+Cuckoo hashing extends the power of two choices by allowing to move
+elements within the table when all chosen positions are already
+occupied (see Section~\ref{sec:pre_cuckoo} for a more elaborated
+explanation).  Cuckoo hashing revitalized research into space
+efficient hash tables.  Probabilistic bounds for the maximum fill
+degree and expected displacement distances are often highly
+non-trivial TODO SOME BOUNDS PAPERS.
 
-Further adaptations to cuckoo hashing include: a de-amortization
+Cuckoo hashing also spawned many variants that further improve its
+performance.  The most notable one being bucket cuckoo hashing
+($d$-ary cuckoo hashing) a variant that groups cells together in
+buckets thus improving cache usage, increasing displacement
+opportunities and maximum fill degree TODO BUNCH OF PAPERS.  There are
+even further variations on bucket cuckoo hashing, like the
+generalization to continuous buckets TODO PAPER.
+
+Further adaptations of cuckoo hashing include:multiple concurrent implementations either powered by bucket
+locking, transactional
+memory~\cite{AlgorithmicImprovementsForFastConcurrentCuckooHashing},
+or fully lock-less~\cite{LockFreeCuckooHashing}; and a de-amortization
 technique that allows a provable worst case analysis of
 insertions~\cite{UsingAQueueToDeAmortizeCuckooHashingInHardware,
-  DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults};
-and multiple concurrent implementations either powered by bucket
-locking, transactional
-memory~\cite{AlgorithmicImprovementsForFastConcurrentCuckooHashing},
-or fully lock-less~\cite{LockFreeCuckooHashing}.
+  DeAmortizedCuckooHashingProvableWorstCasePerformanceAndExperimentalResults}.
 
 Some non-cuckoo space efficient hash tables continue to use linear
-probing variants.  Robin Hood hashing is a technique that was
-originally introduced in 1985~\cite{RobinHoodHashing}. It has regained some popularity in
-recent years -- mainly for its interesting theoretical
-properties and the possibility to reduce the inherent variance of
-linear probing.  Hopscotch~\cite{HopscotchHashing} hashing is a
-technique, that originated in the research of concurrent hash tables, but
-it can also lead to some densely filled hash tables.  For this reason
-it is often used as a technique to achieve space efficient hash
-tables.  This is only possible in a small number of circumstances, as
-the stored bitmaps used in hopscotch hashing create a significant
-memory overhead.
+probing variants.  \emph{Robin Hood hashing} is a technique that was
+originally introduced in 1985~\cite{RobinHoodHashing}. The idea behind
+Robin Hood hashing is to move already stored elements during
+insertions in a way that minimizes the longest possible search
+distance.  Robin Hood hashing has regained some popularity in recent
+years -- mainly for its interesting theoretical properties and the
+possibility to reduce the inherent variance of linear probing.
+\emph{Hopscotch hashing}~\cite{HopscotchHashing} is a technique, that
+originated in the research of concurrent hash tables, but it can also
+manage densely filled hash tables.  Hopscotch hashing also moves
+elements within the table to reduce the maximum search distance.
+Additionally it introduces an acceleration data structure that is
+designed to improve lookup performance especially in cases where the
+searched key is not found.
 
-All these publications show that there is a clear interest in the
-development of space efficient hash tables.  Dynamic hash tables on the
-other hand seem to be considered a solved problem.  There is an
-important paper by Dietzfelbinger at
-al.~\cite{DynamicPerfectHashingUpperAndLowerBounds} that takes on the
-problem of dynamic hash tables from a theoretical perspective.  The
-paper predates cuckoo hashing, and much of the attention for space
-efficient hashing. Thus all memory bounds presented are given without
-tight constant factors.  This lack of development is where we pick up
-and offer fast solutions that support dynamic growing with tight space
-bounds.
+All these publications show that there is a clear interest in
+developing hash tables that can be more and more densely filled.
+Dynamic hash tables on the other hand seem to be considered a solved
+problem.  One paper that takes on the problem of dynamic hash tables
+was written by Dietzfelbinger at
+al.~\cite{DynamicPerfectHashingUpperAndLowerBounds}.  It actually
+predates cuckoo hashing, and much of the attention for space efficient
+hashing.  All memory bounds presented are given without tight constant
+factors.  The lack of implementations and theory about dense dynamic
+hash tables is where we pick up and offer a fast hash table
+implementation, that supports dynamic growing with tight space bounds.
 
 \section{Preliminaries}
-A hash table is a data-structure, which stores key-value-pairs
-($\langle key, data \rangle$) and offers the following functions:
+A hash table is a data-structure for storing key-value-pairs
+($\langle key, data \rangle$) and offers the following functionality:
 \verb~insert~ -- stores a given key-value pair; \verb~find~ -- given a
@@
 
-Throughout this paper, we use $n$ to denote the number of elements and
-$m$ the number of cells ($m > n$) in any given hash table.  We define
-the load factor as $load = n/m$, and we use $\varepsilon = 1-load$ for
-the fraction of available cells.  Tables can usually only work
-efficiently up to a certain load factor.  Above that, operations get
-slower and/or have a possibility to fail.  When implementing a hash
-table one has to decide between storing elements directly in the hash
-table -- \emph{Closed Hashing} -- or storing pointers to elements --
-\emph{Open Hashing}. This has an immediate impact on the necessary
-memory.
-
+Throughout this paper $n$ denotes the number of elements and $m$ the
+number of cells ($m > n$) in any given hash table.  We define the load
+factor as $load = n/m$ and the fraction of empty cells $\varepsilon =
+1-load$.  Tables can usually only work efficiently up to a certain
+load factor.  Above that, operations get slower and/or have a
+possibility to fail.  When implementing a hash table one has to decide
+between storing elements directly in the hash table -- \emph{Closed
+  Hashing} -- or storing pointers to elements -- \emph{Open
+  Hashing}. This has an immediate impact on the necessary memory:
 \begin{align*}
@@
 For large elements ($\gg size(pointer)$), one can use a non-space
-efficient hash table with open hashing, to reduce the relevant memory
+efficient hash table with open hashing to reduce the relevant memory
 factor.  Therefore, we restrict ourselves to the common and more interesting case of
@@
 experiments we use 128bit elements (64bit keys and 64bit values).  In
-these instances open hashing introduces a significant memory
-overhead (at least $\times1.5$), therefore, we only consider closed
-hash tables for the purpose of this paper. For closed hash tables, the
-memory efficiency is directly dependent on the element density.  To
+this case, open hashing introduces a significant memory
+overhead (at least $\times1.5$). For this reason, we only consider closed
+hash tables. For them, the
+memory efficiency is directly dependent on their load.  To
 reach high fill degrees with closed hashing tables, we have to employ
-\emph{open addressing} techniques which means, that elements are not
+\emph{open addressing} techniques which means that elements are not
 stored in predetermined cells, but can be stored in one of several
@@
 this case we define working efficiently as having average insertion
-times close to $O(\frac{1}{\varepsilon})$. This is a natural bound
+times in $O(\frac{1}{\varepsilon})$. This is a natural bound
 since it is the expected number of fully random probes needed to hit
@@
 hashing) cells are the same size as elements. Therefore, being
-$\alpha$ space efficient is the same as operating with a load factor
+$\alpha$-space efficient is the same as operating with a load factor
 of $\alpha^{-1} = n/m$.
@@
 \paragraph*{Dynamic}
-The definition of a space efficient hashing technique given aboce is
-specifically targeted for statically sized hash tables.  We call a
-hash table implementation dynamically $\alpha$-space efficient if an
-instantiated table can grow arbitrarily large over its original
-capacity, while remaining smaller than $\alpha\cdot n_{\max}\cdot
+The definition of a space efficient hashing technique given above is
+specifically targeted for statically sized hash tables.  We call an
+implementation dynamically $\alpha$-space efficient if an instantiated
+table can grow arbitrarily large over its original capacity, while
+remaining smaller than $\alpha\cdot n_{\max}\cdot
 \textit{size}(\textit{element}) + O(1)$ at all times.
@@
 $\textit{m}_{\textit{new}} + \textit{m}_{\textit{old}}$
-cells. Therefore, no normal full table migration can be more then
-2-space efficient.  The only option to perform a full table migration
-with less memory, is to increase the memory in place (see
-Section~\ref{sec:dys_inplace} and
-Section~\ref{sec:exp_competitor}).
+cells. Therefore, a normal full table migration is never better than 2-space
+efficient.  The only option to perform a full table migration with
+less memory is to increase the memory in place (see
+Section~\ref{sec:dys_inplace} and Section~\ref{sec:exp_competitor}).
 
 Similar to static $\alpha$-space efficiency, we will mostly talk about
-the \emph{minimum load factor}, instead of $\alpha$.  We should note,
-that this definition of dynamic $\alpha$-space efficient hash tables
-does not enforce size reduction due to the deletion of elements, but
-it enforces that cells previously used by deleted elements must be reused.
+the \emph{minimum load factor} instead of $\alpha$.  We note that this
+definition of dynamic $\alpha$-space efficient hash tables does not
+enforce size reduction due to the deletion of elements. However it
+enforces that cells previously used by deleted elements must be
+\emph{reused}.
 
@@
 lookup times even in densely filled tables.  The distinguishing
-technique of cuckoo hashing is, that $H$ hash functions ($h_1, ... ,
+technique of cuckoo hashing is that $H$ hash functions ($h_1, ... ,
 h_H$) are used to compute $H$ independent positions. Each element is
@@
 one can often move elements to create space for the current
-element. We call this process displacing elements.
+element. We call this process \emph{displacing} elements.
+
+TODO MAYBE ALREADY GRAPH VIEW
 
@@
 are grouped into buckets of size $B$ ($m/B$ buckets).  Each element
-appointed to one bucket can be stored in any of the bucket's cells .
-Using buckets one can drastically increase the number of displacement
-opportunities.
+assigned to one bucket can be stored in any of the bucket's cells .
+Using buckets one can drastically increase the number of elements that
+can be displaced to make room for a new one, thus decreasing the
+expected length of displacement paths.
 
 \emph{Find} and \emph{remove} operations have a guaranteed constant
-running time. Independent from the tables density, there are $H$
+running time. Independent from the tables density there are $H$
 buckets -- $H\cdot B$ cells -- that have to be searched to find an
@@
 one of those buckets has an empty cell, we store the element in the
-bucket with the most space.  When all buckets are full, then we
-have to move elements around the table, such that a new space becomes
+bucket with the most free space.  When all buckets are full we
+have to move elements within the table such that a free cell becomes
 available.
@@
 directed graph implicitly defined by the hash table.  Each bucket
-defines one node and each element defines pairwise connections between
-the bucket it is stored in to its $H-1$ alternate buckets.  To insert
+corresponds to a node and each element induces an edge between
+the bucket it is stored in and its $H-1$ alternate buckets.  To insert
 an element into the hash table we have to find a path from one of its
 associated buckets, to a bucket, that has space remaining.  Then we
-move elements along this path, to open a space on the initial bucket.
+move elements along this path to make room in the initial bucket.
 The two common techniques to find such paths are \emph{random walks} and
@@
 
-\section{DySECT Space Efficient Growing, by Splitting the Table}
-Our goal is to build an efficiently growing data structure, that
-remains space efficient at all times.  A commonly used growing
-technique is to double the size by migrating all elements
-into a larger new table.  Of course this is
-not memory efficient.  A natural thought is to double
-only part of the data structure.  This idea together with the
-balancing mechanism of cuckoo hashing will achieve the
-described functionality.
+\section{DySECT (Dynamic Space Efficient Cuckoo Table)}
+A commonly used growing technique is to double the size of a hash
+table by migrating all its elements into a table with twice its
+capacity.  This is of course not memory efficient.  The idea behind
+our hash table architecture is to double only parts of the overall
+data-structure.  This increases the space in part of our
+data-structure without changing the rest.  We then use cuckoo
+displacement techniques to make this additional memory reachable from
+other parts of the hash table.
 
-\subsection{Description}
-We use $T$ subtables (shown in TODO FIGREFRENCE) to allow
-differentiated growing between table parts.  The subtables consist of
-buckets that store elements.
-
+\subsection{Overview}
+Our DySECT hash table consists of $T$ subtables (shown in TODO
+FIGREFRENCE), that in turn consist of buckets that store elements.
 Each element has $H$ associated
-buckets similar to cuckoo hashing, these can be in the same, or
+buckets similar to cuckoo hashing which can be in the same or
 in different subtables.  To find a bucket associated with an element $e$,
 we compute $e$'s hash value using the appropriate hash function
-$h_i(e)$, from a hashed key, we have to compute the subtable and the bucket
+$h_i(e)$. The hash is then used to compute the subtable and the bucket
 within that subtable.
@@
 To make this efficient we use powers of two for the number of
-subtables ($T = 2^t$), as well as the individual subtable sizes ($s =
+subtables ($T = 2^t$), as well as for the individual subtable sizes ($s =
 2^x*B$).  Since the number of subtables is constant we can use the
 first $t$ bits from the hashed key, to find the appropriate subtable.
-From the remaining bits we compute the offset into this subtable using
-a bitmask ($h_i(e)\,\texttt{\&}\,(s-1) = h_i(e) \mod s$).
+From the remaining bits we compute the bucket within the subtable using
+a bitmask ($h_i(e)\,\texttt{\&}\,(2^x-1) = h_i(e) \mod 2^x$).
 
@@
 \subsection{Growing}
-We grow one subtable,
-when the table contains enough elements, that the memory constraint
+We grow one subtable by migrating it into a table twice its size
+as soon as the (overall) table contains enough elements such that the memory constraint
 can be kept during the migration.  We migrate subtables in order
-from the first to the last, therefore, no subtable can be more than
+from the first to the last. This ensures that no subtable can be more than
 twice as large as any other.
 
-Assume that all subtables have $s=m/T$ cells. When $\alpha\cdot n >
+Assume that all subtables have the same size ($s=m/T$). When $\alpha\cdot n >
 m+2s$ we can grow the first subtable while keeping the size
@@
 (T+1)\cdot s$ (grow factor $\frac{T+1}{T}$).  Note that all subsequent
-grows migrate one of the smaller tables, until all tables have the
-same size again.  Therefore, each grow until then increase the
-capacity by the same absolute amount (smaller grow factor).
+growing operations migrate one of the smaller tables, until all tables have the
+same size again.  Therefore, each grow until then increases the
+overall capacity by the same absolute amount (smaller overall factor TODO EXPLAIN FURTHER?).
 
-The cost of growing the subtable is amortized by all insertions that
+The cost of growing a subtable is amortized by all insertions that
 happened since the last subtable migration.  There are at least
 $\alpha^{-1} \cdot s = \Omega(s)$ insertions between two migrations.
-One migration takes $\Theta(s)$ time.  Apart from being amortized,
-the migration is efficient in practice because it accesses cells in a
-linear fashion making it really cache efficient.  Even in the target
-table cells are accessed linearly.  Because of the way we assign
-elements to buckets, one original bucket is split into two target
-buckets.  Therefore, no displacements are necessary, since no bucket
-can be overfilled.
+One migration takes $\Theta(s)$ time.  Apart from being amortized, the
+migration is cache efficient since it accesses cells in a linear
+fashion.  Even in the target table cells are accessed linearly.  We
+assign elements to buckets by using bits from their hash value. In the
+grown table we use exactly one more bit than before (double the number
+of buckets).  This ensures, that all elements from one original
+bucket, are split between two buckets in the target table.
+Therefore, no bucket can overflow and no displacements are necessary.
 
@@
 (Section~\ref{sec:cuckoo_graph}), growing a subtable is equivalent to
-splitting each node that represents a bucket within that table. The
-edges (elements) in the implicitly defined subgraph are not doubled,
-therefore, the resulting subgraph becomes more sparse, making it
-easier to insert elements.
+splitting each node that represents a bucket within that table. Since
+the edges (elements) in the implicitly defined subgraph are not
+doubled, the resulting subgraph becomes more sparse, making it easier
+to insert elements.
 
 \subsection{Shrinking}
-In many use cases, shrinking is not necessary.  It worsens performance
-by taking time for the migration and making the remaining table more
-dense.  Furthermore, it is unclear that the freed memory is actually
-necessary for other parts of the application.  If that memory is
-necessary, and shrinking is required, for example, when removed
-elements are reinserted into another data structure and global memory
-bounds are necessary (other data structures grow while the hash table
-shrinks).
+In many use cases automatic shrinking is not necessary.  It worsens
+performance by taking time for the migration and making the remaining
+table more dense.  Therefore, we forgo automatic shrink operations
+whenever enough elements are removed from the table.
 
-Shrinking can work similarly to growing. We replace a subtable with a
-smaller one by migrating elements from one to the other.  During this
-migration we join elements from two buckets into one. Therefore, it is
-possible for a bucket to overfill.  We reinsert these elements at the
-end of the migration.  Obviously, this can only apply to at most half
-the migrated elements.
+If shrinking is necessary it can work similarly to growing. We
+replace a subtable with a smaller one by migrating elements from one
+to the other.  During this migration we join elements from two buckets
+into one. Therefore, it is possible for a bucket to overfill.  We
+reinsert these elements at the end of the migration.  Obviously, this
+can only apply to at most half the migrated elements.
 
-When triggering the size reduction, one has to make sure that the
-migration is amortized. Therefore, a grow operation cannot immediately
-follow a shrink operation.  When shrinking is enabled we propose to
-shrink one subtable when $\alpha\cdot n < m-s'$ elements ($s'$ size of a
-large table, $m_{new} = m_{old} -s'/2$).
+When automatically triggering the size reduction, one has to make sure
+that the migration cost is amortized. Therefore, a grow operation cannot
+immediately follow a shrink operation.  When shrinking is enabled we
+propose to shrink one subtable when $\alpha\cdot n < m-s'$ elements
+($s'$ size of a large table, $m_{new} = m_{old} -s'/2$).
+Alternatively, one could implement a \emph{shrink to size} operation that is
+explicitly called by the user.
 
@@
 
+\subsection{Difficulties for the Analysis of our table structure}
+There are two factors that impact the performance of our dynamic
+table compared to other cuckoo table variants and to other hashing
+solutions in general: \textbf{inhomogeneous table resolution} and
+\textbf{element imbalance}.
+%, and \textbf{population density}.
+Both factors influence the maximum load density and the
+running times in different ways.
+
+\subparagraph*{Imbalance through Inhomogeneous Table Resolution.}
+\label{sec:inhom_res}
+By growing subtables individually we introduce a size imbalance
+between subtables. Large subtables contain more buckets but the number
+of elements hashed to a large subtables is not generally higher than
+the number of elements that are hashed to a small subtable.  This
+makes it difficult to spread elements evenly among buckets.
+Imbalanced bucket fill ratios can lead to longer insertion times.
+
+Assume there are $n$ elements in a hash table with $T$ subtables,
+$j$ of which have size $2s$ the others have size $s$. If elements are
+spread equally among buckets then all small tables have around
+$n/(T+j)$ elements, and the bigger tables have $2n/(T+j)$ elements.
+For each table there are about $Hn/T$ elements associated to one of
+its buckets.  This shows that having more hash functions can lead to a
+better balance.
+
+For two hash functions ($H=2$) and only one grown table ($j=1$) this
+means that $\approx 2n/(T+1)$ elements should be stored in the first table to
+achieve a balanced bucket distribution.  Therefore, nearly all
+elements associated with a bucket in the first table ($\approx 2n/T$)
+have to be stored there.  This is the reason why $H=3$ leads to
+significantly better space efficiency than $H=2$.
+
+\subparagraph*{Imbalance through Size Changes.}
+In addition to the problem of inhomogeneous tables there is an
+inherent balancing problem introduced by resizing subtables. It is
+clear that a newly grown table is not filled as densely as other
+tables.  Since we double the table size, grown tables can only be
+filled to about 50\%.
+
+Assume the global table is filled close to 100\% when the first table
+begins to grow.  Now there is capacity for $s$ new elements but this
+capacity is only in the first table, elements that are not hashed to
+the first table, automatically trigger displacements leading to slow
+insertions.  Notice that repeated insert and erase operations help to
+equalize this imbalance, because elements are more likely inserted
+into the sparser areas, and more likely to be deleted from denser
+areas.
+
+% \subparagraph*{Population Density}
+% Measuring and comparing the performance of space efficient growing
+% tables to their statically sized counterparts is difficult.  Inserting
+% elements into a densely filled table takes longer, than into an empty
+% one.  Therefore, a table that is always densely filled will naturally
+% be slower than one that only fills up towards the end of the
+% execution -- even when ignoring eventual growing costs.
+
 \subsection{Implementation Details}
@@
 functions ($H=3$) and a bucket size of ($B=8$). These values have
-consistently outperformed other options.  We chose to use $T=256$
+consistently outperformed other options.  $T$ is set to 256
 subtables for all our tests.  The number of subtables only plays a
@@
 
-\subsubsection*{Reducing Number Computed Hash Functions}
+\subparagraph*{Reducing Number Computed Hash Functions}
 Evaluating hash functions is expensive, therefore, reducing the number
-of hash functions computed per operation increases the efficiency of
+of hash functions computed per operation increases the performance of
 the table.  The hash function we use computes 64bit hash values
@@
 
-When $H > 2$ we can use \emph{double hashing} to further reduce the number
-of computed hash functions. Double hashing creates an arbitrary amount
+When $H > 2$ we can use \emph{double hashing} TODO CITE to further reduce the number
+of computed hash functions. Double hashing creates an arbitrary number
 of hash values using only two original hash functions $h'$ and
@@
 
-\subsubsection*{Table Architecture (two Variants)}
+\subparagraph*{Table Architecture (two Variants)}
 \label{sec:dys_inplace}
 The hash table can be implemented as a collection of pointers to
-subtables. We have to lookup the corresponding pointer, whenever a
+subtables. We have to lookup the corresponding pointer whenever a
 subtable is accessed.  This does not impact performance much since all
-subtable pointers will usually be cached.
+subtable pointers will be cached -- if the hash table is a performance bottleneck.
 
-Another option, which offers some benefits is to use virtual memory,
+Another option for implementing subtables is to use virtual memory,
 and memory overcommiting.  The idea is that the operating system will
@@
 implement the subtables as sections within one huge allocation.  This
-has the advantage, that the offset for each table can be quickly
-computed, without looking it up from a table.
+has the advantage that the offset for each table can be computed
+quickly, without looking it up from a table.
 
@@
 
-This way, we can also implement in place growing variants of other
-hashing schemes (see Section~\ref{sec:exp_competitors}).  These
-implementations have the same performance of similar implementations
-using full table migration.  But given the correct growing factor,
-they can be space efficient.
-
-\subsection{Difficulties for the Analysis of our table structure}
-There are two factors, that impact the performance of our dynamic
-table compared to other cuckoo table variants and to other hashing
-solutions in general \textbf{inhomogeneous table resolution} and
-\textbf{element imbalance}.
-%, and \textbf{population density}.
-All of these factors influence the maximum load density and the
-running times in different ways.
-
-\subparagraph*{Imbalance through Inhomogeneous Table Resolution}
-\label{sec:inhom_res}
-By growing subtables individually we introduce a size imbalance
-between subtables.  Large tables contain more buckets, but the number
-of elements hashed to the table itself is not dependent on its size,
-therefore, it is difficult to spread elements evenly among buckets.
-Uneven bucket fill ratios can lead to longer insertion times.
-
-If there are $n$ elements in a hash table with $T$ subtables,
-$j$ of which have size $2s$ the others have size $s$. If elements are
-spread equally among buckets then all small tables have around
-$n/(T+j)$ elements, and the bigger tables have $2n/(T+j)$ elements.
-For each table there are about $Hn/T$ elements associated to one of
-its buckets.  This shows that having more hash functions can lead to a
-better balance.
-
-For two hash functions ($H=2$) and only one grown table ($j=1$) this
-means that $\approx 2n/(T+1)$ should be stored in the first table, to
-achieve a balanced bucket distribution.  Therefore, nearly all
-elements associated to a bucket in the first table ($\approx 2n/T$)
-have to be stored there.  This is the reason, why $H=3$ leads to
-significantly better space efficiency than $H=2$.
-
-\subparagraph*{Imbalance through Size Changes}
-In addition to the problem of inhomogeneous tables, there is an
-inherent balancing problem introduced by resizing subtables. It is
-clear, that a newly grown table is not filled as densely as other
-tables.  Since we double the table size, grown tables can only be
-filled to about 50\%.
-
-Assume the global table is filled close to 100\% when the first table
-begins to grow.  now there is the capacity for $s$ new elements, but
-this capacity is only in the first table, elements that are not hashed
-to the first table, automatically trigger displacements leading to
-slow insertions.  Notice that repeated operations help to equalize
-this imbalance, because elements are more likely inserted into the
-less dense areas, and more likely to be deleted from more dense areas.
-
-% \subsubsection*{Population Density}
-% Measuring and comparing the performance of space efficient growing
-% tables to their statically sized counterparts is difficult.  Inserting
-% elements into a densely filled table takes longer, than into an empty
-% one.  Therefore, a table that is always densely filled will naturally
-% be slower than one that only fills up towards the end of the
-% execution -- even when ignoring eventual growing costs.
-
 \section{Experiments}
@@
 these factors are inherent to the combination of growing and space
-efficiency.  To show that our ideas work in practice we use a
-portfolio of both micro-benchmarks and practical experiments.
+efficiency.  To show that our ideas work in practice we use both
+micro-benchmarks and practical experiments.
 
-\subsection{Methodology and Comparison Implementations}
+All reported numbers are averaged by running each experiment five
+times.  All measurements, where more than 100 operations failed are
+omitted, this happens mostly for hopscotch hashing.  The experiments
+were executed on multiple machines.
+TODO...
 
-\subsubsection*{Comparison Implementations}
+The results were very similar, therefore, we only report the numbers
+achieved on the
+
+TODO
+...
+
+\subsection*{Competitors}
 \label{sec:exp_competitors}
@@
 
-\subparagraph*{Linear Probing} is the most common hashing technique outside of
-hashing with chaining.  It is commonly accepted to be inefficient for
-higher load factors because search distances can become large.  In
-practice, this fact is counteracted by its cache-efficiency making
-operations fast in spite of long probing distances.  Linear probing is
-especially slow in \verb~find~ and \verb~erase~ operations, that often have
-guaranteed constant running time in other implementation.
-
-To grow our linear probing table, we use full table migrations.  This
-can be made space efficient, by increasing the capacity in place.
-This is possible due to over allocation (similar to
-Section~\ref{sec:dys_inplace}) and a simple in place migration
-algorithm (APPENDIX OR SECTION?).  For a table that was
-initialized with $m$ cells with a min load factor $1-\varepsilon$ we
-trigger growing once the table is loaded more than
-$1-\frac{\varepsilon}{2}$. We then increase the capacity $m$ to
-$(1-\varepsilon)^{-1}\cdot n$ to remain strictly size-efficient.  Our implementation
-of the migration is highly cache efficient, but all elements have to
-be moved.  This makes repeated migrations inefficient especially for
-small $\varepsilon$.
+\subparagraph*{Linear Probing} is the most common closed hashing
+technique.  It is knownaccepted to be inefficient for higher load
+factors because search distances can become large.  In practice, this
+fact is counteracted by its cache-efficiency making operations fast in
+spite of long probing distances.  Linear probing is especially slow in
+\verb~find~ and \verb~erase~ operations which often have guaranteed
+constant running time in other implementation.
 
@@
 data structure which uses bitmaps to store neighborhood data. This
-data improves find and erase performance.  Many people would expect
-the acceleration data to be the main draw to hopscotch hashing, but it
-is also its greatest demise in the space-efficient setting.  The
-bitmaps introduce a significant memory overhead of neighborhood-size
+data improves \verb~find~ and \verb~erase~ performance.  The
+bitmaps introduce a significant memory overhead of $|$neighborhood$|$
 bits per cell.  Our experiments show that for any reasonable fill
-factor neighborhoods have to be 64bits and larger, this introduces a
+factor neighborhoods have to be 64bits and larger.  This introduces a
 size factor of $\times1.5$ (128bit elements).  We still include
-measurements with hopscotch hashing -- without considering space for
-the neighborhood data -- to show its overall performance tendencies.
-Growing hopscotch tables is done similar to the technique described
-under linear probing.
+measurements with hopscotch hashing \emph{without considering space for
+the neighborhood data} to show its overall performance tendencies.
 
-TODO DESCRIBE 2 CELL BUCKET VARIANT
+Since 64bit neighborhoods failed on many of our tests, we implemented
+a variant with neighborhoods of 128 cells.  In this variant, we only
+store one bitmask per two cells (each bit of the bitmask denotes two
+cells). One could call this variant bucket Hopscotch hashing.
 
-\subparagraph*{Robin Hood Hashing} is another technique based on linear probing. The
-goal of this technique is to improve find performance by minimizing
-the maximum displacement distance.  The drawback of this
-technique is that insertions have to compute the hash of each element
-that is skipped by the insertion.  Migrating a Robin Hood hash table
-is very fast, since the additional hash computations are not
-necessary, and the element order remains the same (sorted by hash value).
+\subparagraph*{Robin Hood Hashing} is a linear probing variant that
+minimizes the maximum search distance.  The drawback of this technique
+is that insertions have to compute the hash of each element that is
+skipped by the insertion.
 
 \subparagraph*{Classic Bucket Cuckoo Hashing} as described in
-Section~\ref{sec:pre_cuckoo} is the basic hashing technique, that
-inspired our hashing technique.  Like our variant it offers constant
-lookup and deletion times.  But it suffers from an inefficient growing
-technique similar to the one used by the linear probing table above.
-Similar to DySECT, $H=3$ and $B=8$ have turned out to be an optimum
-for performance.
+Section~\ref{sec:pre_cuckoo}, similar to DySECT we choose $H=3$ and
+$B=8$ since they have the optimal performance.
 
-\subsubsection*{Hardware}
-We test our implementation on two very different systems, first a
-consumer level desktop computer \ldots{}
+\subparagraph*{Growing our Competitors}
+To grow these competitor implementations, we use in place table
+migrations.  This is possible due to over allocation (similar to
+Section~\ref{sec:dys_inplace}).  We allocate a large piece of memory,
+when the table is created but only a small fraction of this memory is
+initialized.  Increasing the memory of the hash table is similar to
+initializing a new section of the allocated memory.  In addition to
+this allocation technique we implement some simple in place migration
+algorithms (APPENDIX OR SECTION?).  These algorithms are relatively
+cache efficient, and reach a performance similar to other migrations
+(by reinsertion of elements).
 
-\subsection{Influence of Fill Ratio under a Static Table Size}
+For a table that was initialized with a min load factor
+$1-\varepsilon$ we trigger growing once the table is loaded more than
+$1-\frac{\varepsilon}{2}$. We then increase the capacity $m$ to
+$(1-\varepsilon)^{-1}\cdot n$.  Our implementation of the migration is
+highly cache efficient, but all elements have to be moved.  This makes
+repeated migrations inefficient especially for small $\varepsilon$.
+
+\subsection{Influence of Fill Ratio (static table size)}
 \label{sec:exp_eps}
 
-The following test was performed, by initializing a table with
-$m\approx25M$~cells (non-growing). Then elements are inserted, until there
-is a critical amount of failing insertions ($>100$).  At different
-stages, we measure the speed of new insertions
-(Figure~\ref{fig:eps_insert}), and finds (Figure~\ref{fig:eps_find}).
-Performance is measured by performing each operation 1000 times, finds
-are measured using either randomly selected elements from within the table (successful), or by
-searching random elements from key space (unsuccessful).
+The following test was performed by initializing a table with
+$m\approx25\,000\,000$~cells (non-growing). Then elements are inserted
+until there is a critical amount of failing insertions ($>100$).  At
+different stages, we measure the running time of new insertion
+(Figure~\ref{fig:eps_insert}), and find (Figure~\ref{fig:eps_find})
+operations (averaged over 1000 operations).  Finds are measured using
+either randomly selected elements from within the table (successful),
+or by searching random elements from key space (unsuccessful).
 
@@
 
-As one might expect, the performance, of insert operations depends
+As one might expect, the performance of insert operations depends
 highly on the fill degree of the table.  Therefore, we show it
 normalized with $\frac{1}{\varepsilon}$, which is the expected number
-of fully random probes, to find a free cell and thus a natural
-estimate for the running time. We see that -- up to a certain point --
-for all tables the insertion time behaves proportional to
-$\frac{1}{\varepsilon}$, then it begins to decrease uncontrolled.  The
-plot shows, that our table remains stable for longer than linear
-probing variants, but not as long, as for classic cuckoo hashing.
-Hopscotch hashing fails earlier than all other tables, because its
-bitmaps become to small, to store the neighborhood information.
+of fully random probes to find a free cell and thus a natural estimate
+for the running time. We see that -- up to a certain point -- the
+insertion time behaves proportional to $\frac{1}{\varepsilon}$ for all
+tables, then it begins to decrease uncontrolled.  The running time of
+our table remains stable for longer than linear probing variants, but
+not as long, as for classic cuckoo hashing.  Hopscotch hashing fails
+earlier than all other tables, because its bitmaps become to small to
+store the neighborhood information.
 
-DySECT's performance degrades earlier than cuckoo hashing.  One reason
+DySECT's performance degrades earlier than that of cuckoo hashing.  One reason
 for this is that our hash table consists of $T$ subtables with power
-of two sizes and thus does not allow exactly $25M$ cells.  Therefore,
-our table is slightly more full.  Another reason is that to get close
-to $25M$ cells, we have 125 large subtables (twice the size of
+of two sizes and thus does not allow exactly 25\,000\,000 cells.  Therefore,
+our table is slightly more full.  Another reason is that in order to get close
+to 25\,000\,000 cells, we have 125 large subtables (twice the size of
 the others).  This introduces inhomogeneous table resolution (see
-\ref{sec:inhom_res}), which decreases our performance compared to
+Section~\ref{sec:inhom_res}), which decreases our performance compared to
 cuckoo hashing.
 
-Figure~\ref{fig:eps_find} shows that linear probing variants have bad
-performance on find operations on highly filled tables.  It also shows
-that unsuccessful find operations are often very slow when using
-linear probing.  This is where the acceleration data-structure used by
-hopscotch tables can significantly improve performance.  Both of these
-effects, do not affect cuckoo variants like DySECT, that have
+Figure~\ref{fig:eps_find} shows that linear probing variants perform
+badly for find operations on highly filled tables.  It also shows that
+unsuccessful find operations are often very slow when using linear
+probing.  This is where the acceleration data-structure used by
+hopscotch tables can significantly improve performance.  Cuckoo hashing variants like DySECT have
 guaranteed constant running times for all find operations --
-independent of the fill degree.
+independent of their success and the table's filling degree.
 
-\subsection{Constructing a Table with Growing}
+\subsection{Influence of Fill Ratio (dynamic table size)}
 \label{sec:exp_ti}
@@
 %% \end{figure}
-In this test $20M$ elements are inserted into a previously empty
+In this test 20\,000\,000 elements are inserted into an initially empty
 table.  The table is initialized expecting 50\,000 elements, thus
-growing is necessary, to fit all elements.  The tables are configured,
+growing is necessary to fit all elements.  The tables are configured
 to guarantee a load factor of at least
 $1-\varepsilon$. Figure~\ref{fig:ti_insert} shows the performance in
-relation to the load factor.  Insertion times are normalized similar to
-Figure~\ref{fig:eps_insert} (divided by $\frac{1}{\varepsilon}$) to
-make them more readable.
+relation to the load factor.  Insertion times are computed as average of all 20\,000\,000 insertions. They are normalized similar to
+Figure~\ref{fig:eps_insert} (divided by $\frac{1}{\varepsilon}$).
 
@@
 key.  In these cases, only dynamic hash tables can remain space
-efficient even though the number of unique keys is often unknown.  In
+efficient because the number of unique keys is often unknown.  In
 the following test, we take the first block of the common crawl
 data-set (TODO REFERENCE) and compute a word count of the contained
-words.  The chosen block contains around 240M words, with around 20M
+words.  The chosen block contains around 240\,000\,000 words, with around 20\,000\,000
 unique words. For the test, we hash each word and insert it's hash
@@
 The performance results can be seen in Figure~\ref{fig:crawl}.  We do
-not use any normalization since -- on average -- each word is repeated
-12 times, therefore, most inserts will actually behave like successful
-find operations (returning an iterator to the contained element).  Our
-DySECT table manages to dominate the field, similar to the growing
-test in Section~\ref{sec:exp_ti}.  This is a hint to the fact, that
-insert performance is important even in find intensive workloads.
-This insight is intensified in the following test where we mix insert
-and find operations.  (Section~\ref{sec:exp_mix}).
+not use any normalization since each word is repeated 12 times (on
+average).  Therefore, most inserts will actually behave like
+successful find operations.  Our DySECT table manages to dominate the
+field, similar to the growing test in Section~\ref{sec:exp_ti}.  This
+indicates that insert performance can dominate running times even in
+find intensive workloads.  This insight is intensified in the
+following test where we mix \verb~insert~ and \verb~find~ operations.
+(Section~\ref{sec:exp_mix}).
 
@@
 0.95$ (TODO verify).  The test starts a decently filled table
-(15M elements).  On this table we perform 10M operations mixed between inserts and
-find/erase operations.
+(15\,000\,000 elements).  On this table we perform 10\,000\,000
+operations mixed between \verb~insert~ and \verb~find~/\verb~erase~
+operations.
 
